### 2023-11-22 preliminary-summary.tex
update report

\documentclass[hyperref, a4paper]{article}

\usepackage{geometry}
\usepackage{titling}
\usepackage{titlesec}
% No longer needed, since we will use enumitem package
% \usepackage{paralist}
\usepackage{enumitem}
\usepackage{footnote}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{soulutf8}
\usepackage{physics}
\usepackage{tensor}
\usepackage{siunitx}
\usepackage[version=4]{mhchem}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{autobreak}
\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{nameref,zref-xr}
\zxrsetup{toltxlabel}
\usepackage[backend=bibtex,sorting=none]{biblatex}
\addbibresource{arpes.bib}
\usepackage[colorlinks,unicode]{hyperref} % , linkcolor=black, anchorcolor=black, citecolor=black, urlcolor=black, filecolor=black
\usepackage[most]{tcolorbox}
\usepackage{prettyref}

% Page style
\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}
\titlespacing{\paragraph}{0pt}{1pt}{10pt}[20pt]
\setlength{\droptitle}{-5em}

% More compact lists 
\setlist[itemize]{
    itemindent=17pt, 
    leftmargin=1pt,
    listparindent=\parindent,
    parsep=0pt,
}

% Math operators
\DeclareMathOperator{\timeorder}{\mathcal{T}}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\legpoly}{P}
\DeclareMathOperator{\primevalue}{P}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\res}{Res}
\newcommand*{\ii}{\mathrm{i}}
\newcommand*{\ee}{\mathrm{e}}
\newcommand*{\const}{\mathrm{const}}
\newcommand*{\suchthat}{\quad \text{s.t.} \quad}
\newcommand*{\argmin}{\arg\min}
\newcommand*{\argmax}{\arg\max}
\newcommand*{\normalorder}[1]{: #1 :}
\newcommand*{\pair}[1]{\langle #1 \rangle}
\newcommand*{\fd}[1]{\mathcal{D} #1}
\DeclareMathOperator{\bigO}{\mathcal{O}}

% TikZ setting
\usetikzlibrary{arrows,shapes,positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.markings}
\tikzstyle arrowstyle=[scale=1]
\tikzstyle directed=[postaction={decorate,decoration={markings,
    mark=at position .5 with {\arrow[arrowstyle]{stealth}}}}]
\tikzstyle ray=[directed, thick]
\tikzstyle dot=[anchor=base,fill,circle,inner sep=1pt]

% Algorithm setting
% Julia-style code
\SetKwIF{If}{ElseIf}{Else}{if}{}{elseif}{else}{end}
\SetKwFor{For}{for}{}{end}
\SetKwFor{While}{while}{}{end}
\SetKwProg{Function}{function}{}{end}
\SetArgSty{textnormal}

\newcommand*{\concept}[1]{{\textbf{#1}}}

% Embedded codes
\lstset{basicstyle=\ttfamily,
  showstringspaces=false,
  commentstyle=\color{gray},
  keywordstyle=\color{blue}
}

% Reference formatting
\newcommand*{\citesec}[1]{\S~{#1}}
\newcommand*{\citechap}[1]{chap.~{#1}}
\newcommand*{\citefig}[1]{Fig.~{#1}}
\newcommand*{\citetable}[1]{Table~{#1}}
\newcommand*{\citepage}[1]{pp.~{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newcommand*{\term}[1]{\textit{#1}}

% Color boxes
\tcbuselibrary{skins, breakable, theorems}

\newtcbtheorem{infobox}{Box}{
    enhanced,
    boxrule=0pt,
    colback=blue!5,
    colframe=blue!5,
    coltitle=blue!50,
    borderline west={4pt}{0pt}{blue!65},
    sharp corners,
    fonttitle=\bfseries, 
    breakable,
    before upper={\parindent15pt\noindent}}{box}
\newtcbtheorem[use counter from=infobox]{theorybox}{Box}{
    enhanced,
    boxrule=0pt,
    colback=orange!5, 
    colframe=orange!5, 
    coltitle=orange!50,
    borderline west={4pt}{0pt}{orange!65},
    sharp corners,
    fonttitle=\bfseries, 
    breakable,
    before upper={\parindent15pt\noindent}}{box}
\newtcbtheorem[use counter from=infobox]{learnbox}{Box}{
    enhanced,
    boxrule=0pt,
    colback=green!5,
    colframe=green!5,
    coltitle=green!50,
    borderline west={4pt}{0pt}{green!65},
    sharp corners,
    fonttitle=\bfseries, 
    breakable,
    before upper={\parindent15pt\noindent}}{box}

\newenvironment{shelldisplay}{\begin{lstlisting}}{\end{lstlisting}}

\newcommand*{\kB}{k_{\text{B}}}
\newcommand*{\muB}{\mu_{\text{B}}}
\newcommand*{\efermi}{E_{\text{F}}}
\newcommand*{\pfermi}{p_{\text{F}}}
\newcommand*{\vfermi}{v_{\text{F}}}
\newcommand*{\sA}{\text{A}}
\newcommand*{\sB}{\text{B}}
\newcommand*{\Tc}{T_{\text{c}}}
\newcommand*{\hethree}{$^3$He}
\newcommand*{\hefour}{$^4$He}
\newcommand{\epsr}{\epsilon_{\text{r}}}
\newcommand{\chie}{\chi_{\text{e}}}
\newcommand{\Efreq}{\tilde{\vb*{E}}}
\newcommand{\Dfreq}{\tilde{\vb*{D}}}
\newcommand{\Pfreq}{\tilde{\vb*{P}}}

\title{Preliminary results on trion signature in time-resolved ARPES}
\author{}
\date{}

\begin{document}

\maketitle

When the temporal separation between the pump and the probe is long enough,
the effect of the pump is essentially preparing an initial state 
with various excitations, 
which are then detected by the probe.
The probe drives the electron part(s) of an excitation out the material, 
which then contributes to the total ARPES intensity;
since the electrons in the excitations always reside on conduction bands, 
they leave signatures in the resulting ARPES spectrum 
that are absent in conventional ARPES.
The probability to detect an electron with momentum $\vb{k}$ 
from an excitation $\ket*{\Psi_n}$ at time $t$
is given as \cite{rustagi2018photoemission}
\begin{equation}
    P_{\vb{k}}(\omega, t) = \sum_{c, \vb{k}'}
    { \abs*{M^{fc}_{\vb{k} \vb{k}'}}^2 }
    \int_{t_0}^t \dd{t_1} \int_{t_0}^t \dd{t_2}
    { \ee^{-\ii (E_n - \omega ) (t_1 - t_2)}} 
    { \mel{\Psi_n(t_0)}{c^\dagger_{\vb{k}'} 
    U(t_2, t_1) c_{\vb{k}'}}{\Psi_n(t_0)}}
    { s(t_1) s(t_2)},
    \label{eq:arpes-pure-state}
\end{equation}
where $t_0$ is the time the probe starts,
$\omega$ is the energy of the out-coming electron 
plus the work function minus the photon energy, 
$E_n$ is the energy of $\ket*{\Psi_n(t)}$,
$c$ is a band index in the material, 
$c_{\vb{k}'}$ refers to the annihilation operator of that mode, 
$M_{\vb{k} \vb{k}'}^{fc}$ is the dipole transition matrix 
between the $(c, \vb{k}')$ band electron state in the material and 
the outgoing final state $(f, \vb{k})$,
$U(t_2, t_1)$ is the many-body time evolution operator in the material,
and $s(t)$ is the shape of the probe pulse.
The time evolution operator carries information regarding 
the oscillation of the state of the system after one electron is kicked out,
leading to a broadened form of energy conservation 
after the integrations over $t_1$ and $t_2$ are finished
that determines the shape of the ARPES signature.
It is possible that the final state of pumping is a mixed state 
of several excitations, 
and in this case \eqref{eq:arpes-pure-state} needs to be 
averaged over the probabilistic distribution of the excitations,
and the excitation signatures are overlaid to each other.

In the case of two-dimensional materials, $\vb{k}_\parallel = \vb{k}'$.
Within the framework of the two-band model, 
it has been shown~\cite{rustagi2018photoemission} that 
the signature of an exciton has the dispersion relation of
\begin{equation}
    \omega = E_{S \vb{Q}} + \epsilon_{v \vb{k}_\parallel - \vb{Q}},
\end{equation}
where $E_{S \vb{Q}}$ is the single exciton energy, 
i.e. band gap plus the negative binding energy.
In the case of a positive trion, 
there is no single, definite dispersion relation that determines the shape of the ARPES signature;
assuming that after one electron is driven out, 
scatterings between the two remaining holes are negligible, 
the dispersion of the trion ARPES signature is determined by the following family of curves 
\begin{equation}
    \omega = E_{S \vb{P}} + \epsilon_{v \vb{k}_{\text{h1}}} + \epsilon_{v \vb{k}_{\text{h2}}}, \quad 
    \vb{k}_\parallel - \vb{k}_{\text{h1}} - \vb{k}_{\text{h2}} = \vb{P}.
\end{equation}
In \prettyref{fig:arpes-exciton-trion}, it could be observed that 
in the zero-temperature case,
even when a trion mode and an exciton mode with the same energy,
they still result in different ARPES signatures:
the slopes and curvatures of the ARPES signature are different, 
and when the total momentum of the excitation is given, 
the position shifts of the signature on the ARPES heatmap are also different.

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{images/exciton-trion-comparison.pdf}
    \caption{Comparison between exciton signatures and trion signatures.}
    \label{fig:arpes-exciton-trion}
\end{figure}

\printbibliography

\end{document}

### 2023-11-22 src/exciton.jl

@@ -119,8 +119,8 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
     
     ylims!(p, (-0.8, 1.8))
     xlims!(p, (-0.8, 0.8)) 
-    xlabel!(p, raw"$k$ / Å")
-    ylabel!(p, raw"$\omega$ / eV")
+    xlabel!(p, raw"$k$ (Å)")
+    ylabel!(p, raw"$\omega$ (eV)")
     
     cb = heatmap(
         [1 0; 0 1], 
@@ -185,18 +185,18 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
 
     k_points_near_valley = LinRange(-0.5, 0.5, 100)
     plot!(p, k_points_near_valley, ϵ_c.(k_points_near_valley) .+ E_B,
-        label = raw"Varying $\mathbf{Q}_\mathrm{X}$, max $\left|A_{\mathbf{k}}^{S \mathbf{Q}}\right|^2$",
+        label = raw"Varying $\mathbf{Q}$, max $\left|A_{\mathbf{k}}^{S \mathbf{Q}}\right|^2$",
         linestyle = :dot,
         c = colorant"firebrick2")
     plot!(p, k_points_near_valley, ϵ_v.(k_points_near_valley .- Q) .+ E_SQ(Q),
-        label = raw"fixed $\mathbf{Q}_\mathrm{X}$, varying $\mathbf{k}$",
+        label = raw"fixed $\mathbf{Q}$, varying $\mathbf{k}$",
         linestyle = :dot,
         c = colorant"deepskyblue2")
     
     ylims!(p, (-0.8, 1.8))
     xlims!(p, (-0.8, 0.8)) 
-    xlabel!(p, raw"$k$ / Å")
-    ylabel!(p, raw"$\omega$ / eV")
+    xlabel!(p, raw"$k$ (Å)")
+    ylabel!(p, raw"$\omega$ (eV)")
     
     cb = heatmap(
         [1 0; 0 1], 

### 2023-11-22 src/trion-dispersion.jl

@@ -187,26 +187,29 @@ end
         eV * norm.(k_points_near_valley).^2 * (m_T / (2 * m_c^2)) 
         .- eV * norm.(k_points_near_valley).^2 * (m_v / m_c^2)
         .+ E_B .+ E_g,
-        label = raw"Varying $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = \mathbf{k}_2 = 0$",
+        #label = raw"Varying $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = \mathbf{k}_2 = 0$",
+        label = raw"Varying $\mathbf{P}$, $\mathbf{k}_{\mathrm{h1}} = \mathbf{k}_{\mathrm{h2}} = - \frac{m_{\mathrm{h}}}{M} \mathbf{P}$",
         linestyle = :dot,
         c = colorant"firebrick2")
     # Dispersion relation when P_T is fixed to a constant 
     plot!(p, kx_near_valley, 
         ϵ_v.([(m_c + m_v) / m_T * Q_point] .- k_points_near_valley) 
         .- eV * norm(Q_point)^2 * (m_v / 2m_T^2) .+ eV * norm(Q_point)^2 / 2m_T .+ E_B .+ E_g,
-        label = raw"fixed $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = 0$ or $\mathbf{k}_2 = 0$",
+        #label = raw"fixed $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = 0$ or $\mathbf{k}_2 = 0$",
+        label = raw"fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{h1}}$ or $\mathbf{k}_{\mathrm{h2}}$ is $- \frac{m_{\mathrm{h}}}{M} \mathbf{P} $",
         linestyle = :dot, 
         c = colorant"aqua")
     plot!(p, kx_near_valley, 
         2 * ϵ_v.(([Q_point] .- k_points_near_valley) ./ 2) 
         .+ eV * norm(Q_point)^2 / 2m_T .+ E_B .+ E_g,
-        label = raw"fixed $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = \mathbf{k}_2 $",
+        label = raw"fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{h1}} = \mathbf{k}_{\mathrm{h2}} $",
[…]

### 2023-11-28 preliminary-summary.tex
update derivation of trion Hamiltonian

@@
 where $E_{S \vb{Q}}$ is the single exciton energy, 
-i.e. band gap plus the negative binding energy.
+i.e. band gap plus the negative binding energy, 
+and $\epsilon_{v \vb{k}} < 0$ is the valence band energy from the Fermi energy.
 In the case of a positive trion, 
@@
     \includegraphics[width=0.9\textwidth]{images/exciton-trion-comparison.pdf}
-    \caption{Comparison between exciton signatures and trion signatures.}
+    \caption{Comparison between exciton signatures and trion signatures. 
+    (a), (b) Exciton ARPES signatures at $Q_x = \text{\SI{0}{\angstrom} and \SI{0.5}{\angstrom}}$.
+    (c), (d) Exciton ARPES signatures at $Q_x = \text{\SI{0}{\angstrom} and \SI{0.5}{\angstrom}}$. 
+    The binding energy of the trion mode is manually shifted to
+    the exciton binding energy in \cite{rustagi2018photoemission}, 
+    but significant differences can still be spotted 
+    between exciton and trion signatures.}
     \label{fig:arpes-exciton-trion}

### 2023-12-29 paper/main.tex
update trion code

\documentclass[hyperref, a4paper]{article}

\usepackage{textgreek}
\usepackage{geometry}
\usepackage{titling}
\usepackage{titlesec}
\usepackage{footnote}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{physics}
\usepackage{tensor}
\usepackage{siunitx}
\usepackage[version=4]{mhchem}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{autobreak}
\usepackage[colorlinks,unicode]{hyperref} % , linkcolor=black, anchorcolor=black, citecolor=black, urlcolor=black, filecolor=black
\usepackage{xurl}
\usepackage[most]{tcolorbox}
\usepackage[backend=bibtex,sorting=none,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{trion.bib}
\addbibresource{arpes.bib}

\usepackage{prettyref}

% Page style; to be removed when this article is placed in another template
\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}
\titlespacing{\paragraph}{0pt}{1pt}{10pt}[20pt]
\setlength{\droptitle}{-5em}

% Math operators
\DeclareMathOperator{\timeorder}{\mathcal{T}}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\legpoly}{P}
\DeclareMathOperator{\primevalue}{P}
\DeclareMathOperator{\sgn}{sgn}
\newcommand*{\ii}{\mathrm{i}}
\newcommand*{\ee}{\mathrm{e}}
\newcommand*{\const}{\mathrm{const}}
\newcommand*{\suchthat}{\quad \text{s.t.} \quad}
\newcommand*{\argmin}{\arg\min}
\newcommand*{\argmax}{\arg\max}
\newcommand*{\normalorder}[1]{: #1 :}
\newcommand*{\pair}[1]{\langle #1 \rangle}
\newcommand*{\fd}[1]{\mathcal{D} #1}
\DeclareMathOperator{\bigO}{\mathcal{O}}

% Embedded codes
\lstdefinestyle{console}{
    basicstyle=\footnotesize\ttfamily,
    breaklines=true,
    postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
}

% Reference formatting
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{tbl}{Table~\ref{#1}}

% TiKZ settings
\usetikzlibrary{calc}
\tikzset{every picture/.style={line width=0.3pt}}

% Displaying chemical formula in bookmarkers

\pdfstringdefDisableCommands{%
  \def\\{}%
  \def\ce#1{<#1>}%
}

\pdfstringdefDisableCommands{%
  \def\texttt#1{<#1>}%
  \def\mathbb#1{#1}%
}
\pdfstringdefDisableCommands{\def\eqref#1{(\ref{#1})}}

\makeatletter
\pdfstringdefDisableCommands{\let\HyPsd@CatcodeWarning\@gobble}
\makeatother

\newcommand{\citetime}[1]{item \ref{#1} in \prettyref{sec:timeline}}
\newcommand{\address}[1]{\href{#1}{\url{#1}}}
\newcommand{\shortcode}[1]{\texttt{#1}}

\DeclareSIUnit{\au}{a.u.}

\lstset{style = console}

% Labels 
\newcommand*{\ke}{\vb{k}_{\text{e}}}
\newcommand*{\khi}[1]{\vb{k}_{\text{h#1}}}
\newcommand*{\re}{\vb{r}_{\text{e}}}
\newcommand*{\rhi}[1]{\vb{r}_{\text{h#1}}}
\newcommand*{\me}{m_{\text{e}}}
\newcommand*{\mh}{m_{\text{h}}}
\newcommand*{\Eg}{E_{\text{g}}}

\title{Trion ARPES signature}
\author{Jinyuan Wu, someone else, Diana Y. Qiu}

\begin{document}

\maketitle

\section{ARPES theory}

When the probe is much weaker than the pump 
and does not itself influence the electronic structure,
the most generic theory of driven ARPES is based on the  
pumped single-electron lesser Green function, 
to which the pump adds a self-energy correction,
\cite{freericks2009theoretical,schuler2021theory};
the theory reduced to the usual Fermi golden rule 
when the pump itself is weak \cite{sobota2021angle,freericks2021two}.
When the pump pulse is strong at its peak but has faded away at the probing time, 
the main effect of pumping is preparing an initial state, 
and the ARPES intensity can be evaluated from 
pure-state first-order time dependent perturbation theory,
which gives a Fermi golden rule-like final result \cite{rustagi2018photoemission}.
Here we generalize the methodology in \cite{rustagi2018photoemission} 
to all possible exciton states caused by the pump and not just excitons.

\section{Two-band model for trion}

In this work, we generalize the two-band model approach to trions \cite{chang2021variationally}
to the indirect band gap case.
We start with the following example.
\begin{equation}
    H = \frac{(\ke - \vb{w})^2}{2\me} + \Eg 
    + \frac{\khi{1}^2}{2\mh} + \frac{\khi{2}^2}{2\mh} 
    + V(\rhi{1} - \rhi{2}) - V(\re - \rhi{1}) - V(\re - \rhi{2}).
    \label{eq:two-band-w}
\end{equation}
Here $\khi{i}$ is the crystal momentum of the quantum state containing hole $i$ only, 
which is the opposite of the momentum of the corresponding electron state of the hole;
similarly, $\vb{k}^2/2\mh$ is the opposite of the dispersion relation of the valence band top,
and therefore $\mh > 0$.

It can be easily seen that the total momentum of the trion 
\begin{equation}
    \vb{P} = \ke + \khi{1} + \khi{2}
\end{equation}
is conserved, and we are to rewrite the Hamiltonian in terms of 
the total momentum, which should only appear in the total kinetic energy term,
and two internal degrees of freedom.
Following the procedure in \cite{chang2021variationally},
we define 
\begin{equation}
    \vb{r}_1 = \rhi{1} - \re, \quad \vb{r}_2 = \rhi{2} - \re 
\end{equation}
as the internal degrees of freedom.
To separate the Hamiltonian into terms about $\vb{P}$ and terms about the internal degrees of freedom,
we need to find the canonical coordinate corresponding to $\vb{P}$
and the canonical momenta corresponding to $\vb{r}_{1, 2}$.

Regarding the canonical coordinate corresponding to $\vb{P}$, 
intuitively we may choose the center of mass 
\begin{equation}
    \vb{R} = \frac{\me \re + \mh \rhi{1} + \mh \rhi{2}}{M}, \quad M = \me + 2 \mh
\end{equation}
as the canonical coordinate corresponding to $\vb{P}$;
this correctness of this choice can be directly verified by  
writing $\vb{R}, \vb{r}_1, \vb{r}_2$ as functions of $\re, \rhi{1}, \rhi{2}$
and verifying that $\partial_{\vb{R}}$ is the sum of 
$\partial_{\re}$, $\partial_{\rhi{1}}$ and $\partial_{\rhi{2}}$,
and hence $\vb{P} = - \ii \hbar \partial_{\vb{R}}$ 
is indeed the total momentum.

We can define the canonical momenta of $\vb{r}_{1,2}$, 
hereafter referred to as $\vb{k}_{1,2}$, 
as $- \ii \hbar_{\vb{r}_{1, 2}}$;
in this way 
\begin{equation}
    \ke = \frac{\me}{M} \vb{P} - \vb{k}_1 - \vb{k}_2, \quad 
    \khi{1,2} = \vb{k}_{1, 2} + \frac{\mh}{M} \vb{P},
\end{equation}
and the kinetic energy part of the \eqref{eq:two-band-w} now reads 
\begin{equation}
    H_{\text{kin}} = \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg 
    + \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh}
    - \frac{\vb{w}^2}{2M} + \frac{(\vb{k}_1 + \vb{k}_2 + \vb{w})^2}{2\me}.
\end{equation}
We however want to shift the definitions of $\vb{k}$ 
so that the kinetic part resembles the kinetic energy in a direct band gap system.
By redefining 
\begin{equation}
    \vb{k}_{1,2} = - \ii \hbar \vb{r}_{1, 2} + \frac{\vb{w}}{1 + \me / \mu}, \quad 
    \mu = \frac{\me \mh}{\me + \mh},
\end{equation}
and therefore 
\begin{equation}
    \ke = \frac{\me}{M} \vb{P} + \frac{2 \mh \vb{w}}{M} - \vb{k}_1 - \vb{k}_2, \quad 
    \khi{1,2} = \vb{k}_{1,2} - \frac{\mh \vb{w}}{M} + \frac{\mh}{M} \vb{P},
\end{equation}
the total Hamiltonian now reads
\begin{equation}
    H =  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg 
    + \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
    + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2).
\end{equation}
The redefinition of $\vb{k}_{1, 2}$ does not change the commutation relation 
between $\vb{r}_{1, 2}$ and $\vb{k}_{1, 2}$.
We find that the internal part of the Hamiltonian has the same form 
as the Hamiltonian in \cite{chang2021variationally}.

\section{Variational wave function}

The spin part is in a singlet state, 
and therefore the orbital part of the wave function is symmetric.

An immediate consequence of this fact is that
since at the ground state 
\begin{equation}
    \ke = \frac{\me}{M} \vb{P} + \frac{2 \mh}{M} \vb{w}.
\end{equation}

The same procedure can be applied to other configurations of peaks and valleys.

\section{ARPES signatures}

When $\vb{k}_1$ is set to zero, we find 
\begin{equation}
    \khi{2} = \frac{\me + \mh}{M} \vb{P} + \frac{\mh}{M} \vb{w} - \ke,
\end{equation}

and when $\vb{k}_1 = \vb{k}_2$, we have

\printbibliography

\end{document}

### 2023-12-29 src/arpes.jl

@@ -1,0 +1,35 @@
+using Plots
+using LinearAlgebra
+using Colors
+using ColorSchemes
+using LaTeXStrings
+
+function transparency_gradience(color; steps = 100)
+    α = alpha(color)
+    r = red(color)
+    b = blue(color)
+    g = green(color)
+    
+    map(LinRange(0.0, α, steps)) do α_new
+        RGBA(r, g, b, α_new)
+    end
+end
+
+function white_gradience(color; steps = 100)
+    ColorScheme([colorant"white", color])[LinRange(0, 1, steps)]
+end
+
+function arpes_colormap(gradient)
+    rainbow = cgrad(:rainbow)
+    starting_region = gradient(first(rainbow), steps = 50) 
+    middle_region = rainbow[LinRange(0, 1, 500)]
+    end_region = reverse(gradient(last(rainbow), steps = 400))
+    
+    PlotUtils.ContinuousColorGradient([starting_region..., middle_region..., end_region...])
+end
+
+function custom_colorbar(p, cb)
+    l = @layout [a{0.95w} b]
+    plot(p, cb, layout = l)
+end
+

### 2023-12-29 src/trion-solver.jl

@@ -1,0 +1,73 @@
+using StaticArrays
+
+const fs = 1.5193    
+const inv_eV = 3.80998
+
+struct IndirectTwoBandModel2D
+    m_e :: Float64     # Conduction band effective mass
+    m_h :: Float64     # Valance band effective mass (positive)
+    E_g :: Float64     # Single particle band gap
+    w   :: SVector{2, Float64} # k_{valley} - k_{peak}
+end
+
+struct Dielectric2D 
+    ϵ :: Float64
+    # TODO
+end
+
+# The trion band index S and momentum Q should be pre-determined
+function single_trion_arpes_signature_def(
+    ham::IndirectTwoBandModel2D, 
+    E_B,                                 # Binding energy in eV
+    A_SQ_k1k2,                           # The internal part of the trion wave function  
+    P_T::SVector{2, Float64},            # Total momentum of the trion  
+    k_points,       # The 1BZ sampling 
+    ω_points,       # The frequency sampling, in eV 
+    k_points_ARPES, # k points included in the ARPES plotting
+    broadening,     # Broadening function
+)
+
+    m_e = ham.m_e
+    m_h = ham.m_h
+    E_g = ham.E_g
+    M = 2m_h + m_e
+    w = ham.w
+
+    E_S_PT = inv_eV * norm(P_T - w)^2 / 2M
+
+    heatmap_points = Iterators.product(k_points_ARPES, ω_points) 
+
+    A_kω = map(heatmap_points) do (k_e, ω)
+        sum(k_points) do k_h1
+            k_h2 = P_T - k_e - k_h1
+            ϵ_v_kh1 = - inv_eV * norm(k_h1)^2 / 2m_h
+            ϵ_v_kh2 = - inv_eV * norm(k_h2)^2 / 2m_h
+
+            δ_factor = broadening(ω - E_S_PT - ϵ_v_kh1 - ϵ_v_kh2 - E_B - E_g)
+            
+            k_1 = k_h1 + m_h / M * w - m_h / M * P_T
+            k_2 = k_h2 + m_h / M * w - m_h / M * P_T
+            trion_structure_factor = abs(A_SQ_k1k2(k_1, k_2))^2
+            
+            δ_factor * trion_structure_factor
+        end
+    end
+    
+    A_kω
+end
+
[…]

### 2023-12-29 src/trion.jl

@@ -1,0 +1,108 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion 
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 10,
+    broadening(x) = exp(- σ^2 * x^2),
+    kx_points = LinRange(-1.2, 1.2, 100),
+    ky_points = LinRange(-1.2, 1.2, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.5,
+    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+
+    A_kω_Q = single_trion_arpes_signature(
+        ham, E_B, 
+        ground_state_1s(ham, dielectric), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    p = heatmap(kx_points, ω_points, A_kω_Q', 
+        c = arpes_colormap(transparency_gradience),
+        colorbar = false, 
+        legend_foreground_color = :transparent,
+        legend = :topleft,
+        dpi = 500, 
+        grid = false, 
+        framestyle = :box,
+    )
+
+    plot!(p, kx_points, ϵ_v.(kx_points), 
+        label = "",
[…]

### 2023-12-29 src/trion-arpes-heatmap-def.jl
optimization

@@ -1,0 +1,108 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion, slow version 
+@profview let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 10,
+    broadening(x) = exp(- σ^2 * x^2),
+    kx_points = LinRange(-1.2, 1.2, 100),
+    ky_points = LinRange(-1.2, 1.2, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.5,
+    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+
+    A_kω_Q = single_trion_arpes_signature_def(
+        ham, E_B, 
+        ground_state_1s(ham, dielectric), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    p = heatmap(kx_points, ω_points, A_kω_Q', 
+        c = arpes_colormap(transparency_gradience),
+        colorbar = false, 
+        legend_foreground_color = :transparent,
+        legend = :topleft,
+        dpi = 500, 
+        grid = false, 
+        framestyle = :box,
+    )
+
+    plot!(p, kx_points, ϵ_v.(kx_points), 
+        label = "",
[…]

### 2023-12-29 src/trion-arpes-heatmap-optimized.jl

@@ -2,8 +2,8 @@ using ProgressMeter
 include("trion-solver.jl")
 include("arpes.jl")
 
-# Finite momentum trion 
-@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+# Finite momentum trion, slow version 
+@profview let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_h = 0.8184,
     m_e = 0.4794,
     E_g = 0.7830,
@@ -11,7 +11,6 @@ include("arpes.jl")
     E_B = -0.1,
     w = 0.7,
     β = 10,
-    broadening(x) = exp(- σ^2 * x^2),
     kx_points = LinRange(-1.2, 1.2, 100),
     ky_points = LinRange(-1.2, 1.2, 100),
     dkx = step(kx_points), 
@@ -31,8 +30,9 @@ include("arpes.jl")
 
     ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
     dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = single_trion_arpes_signature(
+    A_kω_Q = single_trion_arpes_signature_thread(
         ham, E_B, 
         ground_state_1s(ham, dielectric), 
         Q_point, 
@@ -71,7 +71,7 @@ include("arpes.jl")
         .- inv_eV * norm.(k_points_near_valley .- [[w, 0.0]]).^2 * (m_h / m_e^2)
         .+ E_B .+ E_g,
         #label = raw"Varying $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = \mathbf{k}_2 = 0$",
-        label = raw"Varying $\mathbf{P}$, $\mathbf{k}_{\mathrm{h1}} = \mathbf{k}_{\mathrm{h2}} = - \frac{m_{\mathrm{h}}}{M} \mathbf{P}$",
+        label = raw"Varying $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}} = \mathbf{k}_{\mathrm{2}} = 0$",
         linestyle = :dot,
         c = colorant"firebrick2")
[…]

### 2023-12-29 src/trion-arpes-optimization-test.jl

@@ -1,0 +1,58 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion, slow version 
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 10,
+    kx_points = LinRange(-1.2, 1.2, 100),
+    ky_points = LinRange(-1.2, 1.2, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.5,
+    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
+
+    A_kω_Q = single_trion_arpes_signature_thread(
+        ham, E_B, 
+        ground_state_1s(ham, dielectric), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    A_kω_Q_def = single_trion_arpes_signature_def(
+        ham, E_B, 
+        ground_state_1s(ham, dielectric), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    sum(abs.(A_kω_Q_def - A_kω_Q)) 
+end

### 2023-12-29 src/trion-solver.jl

@@ -1,4 +1,5 @@
 using StaticArrays
+using ThreadTools
 
 const fs = 1.5193    
 const inv_eV = 3.80998
@@ -15,11 +16,12 @@ struct Dielectric2D
     # TODO
 end
 
-# The trion band index S and momentum Q should be pre-determined
+
 function single_trion_arpes_signature_def(
     ham::IndirectTwoBandModel2D, 
     E_B,                                 # Binding energy in eV
     A_SQ_k1k2,                           # The internal part of the trion wave function  
+                                         # The trion band index S and momentum Q should be pre-determined
     P_T::SVector{2, Float64},            # Total momentum of the trion  
     k_points,       # The 1BZ sampling 
     ω_points,       # The frequency sampling, in eV 
@@ -58,6 +60,62 @@ end
 
 single_trion_arpes_signature = single_trion_arpes_signature_def
 
+function single_trion_arpes_signature_thread(
+    ham::IndirectTwoBandModel2D, 
+    E_B,                                 # Binding energy in eV
+    A_SQ_k1k2,                           # The internal part of the trion wave function  
+    P_T::SVector{2, Float64},            # Total momentum of the trion  
+    k_points,       # The 1BZ sampling 
+    ω_points,       # The frequency sampling, in eV 
+    k_points_ARPES, # k points included in the ARPES plotting
+    broadening,     # Broadening function
+)
+
+    m_e = ham.m_e
+    m_h = ham.m_h
+    E_g = ham.E_g
+    M = 2m_h + m_e
+    w = ham.w
+
+    E_S_PT = inv_eV * (P_T - w)' * (P_T - w) / 2M
+
+    heatmap_points = Iterators.product(k_points_ARPES, ω_points) 
+    
+    A_kω = tmap(heatmap_points) do (k_e, ω)
+        sum(k_points) do k_h1
+            k_h2 = P_T - k_e - k_h1
+            ϵ_v_kh1 = - inv_eV * k_h1' * k_h1 / 2m_h
[…]

### 2023-12-30 src/trion-arpes-heatmap-optimized-one-Q.jl
finite temperature case

@@ -1,0 +1,108 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion, slow version 
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 10,
+    kx_points = LinRange(-1.2, 1.2, 100),
+    ky_points = LinRange(-1.2, 1.2, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.7,
+    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
+
+    A_kω_Q = single_trion_arpes_signature_thread(
+        ham, E_B, 
+        ground_state_1s(ham, dielectric), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    p = heatmap(kx_points, ω_points, A_kω_Q', 
+        c = arpes_colormap(transparency_gradience),
+        colorbar = false, 
+        legend_foreground_color = :transparent,
+        legend = :topleft,
+        dpi = 500, 
+        grid = false, 
+        framestyle = :box,
+    )
+
+    plot!(p, kx_points, ϵ_v.(kx_points), 
+        label = "",
[…]

### 2023-12-30 src/trion-arpes-heatmap-optimized-thermalized.jl

@@ -3,7 +3,7 @@ include("trion-solver.jl")
 include("arpes.jl")
 
 # Finite momentum trion, slow version 
-@profview let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_h = 0.8184,
     m_e = 0.4794,
     E_g = 0.7830,
@@ -21,7 +21,11 @@ include("arpes.jl")
     ),
     ω_points = LinRange(0, 1.5, 500),
     P_Tx = 0.5,
-    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+    Qx_points =  LinRange(w - 0.1, w + 0.1, 10),
+    Qy_points =  LinRange(- 0.1, 0.1, 10),
+    Q_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(Qx_points, Qy_points))
+    ) 
 
     ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
     ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
@@ -32,18 +36,26 @@ include("arpes.jl")
     dielectric = Dielectric2D(ϵ)
     broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = single_trion_arpes_signature_thread(
-        ham, E_B, 
-        ground_state_1s(ham, dielectric), 
-        Q_point, 
-        k_points, ω_points, 
-        map(kx -> SVector{2, Float64}([kx, 0.0]), 
-            kx_points
-        ),
-        broadening
-    )     
+    Z_partition = sum(Q_points) do Q_point
+        exp(-β * (Q_point - [w, 0.0])' * (Q_point - [w, 0.0]) / 2M)
+    end
+    progress = Progress(length(Q_points)) 
+    A_kω_β = sum(Q_points) do Q_point
+        next!(progress)
+        A_kω_Q = single_trion_arpes_signature_thread(
+            ham, E_B, 
+            ground_state_1s(ham, dielectric), 
+            Q_point, 
+            k_points, ω_points, 
[…]

### 2023-12-30 src/trion-arpes-optimization-test.jl

@@ -20,7 +20,7 @@ include("arpes.jl")
         collect(Iterators.product(kx_points, ky_points))
     ),
     ω_points = LinRange(0, 1.5, 500),
-    P_Tx = 0.5,
+    P_Tx = 0.8,
     Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
 
     ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
@@ -32,7 +32,7 @@ include("arpes.jl")
     dielectric = Dielectric2D(ϵ)
     broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = single_trion_arpes_signature_thread(
+    @time A_kω_Q = single_trion_arpes_signature_thread(
         ham, E_B, 
         ground_state_1s(ham, dielectric), 
         Q_point, 
@@ -43,7 +43,7 @@ include("arpes.jl")
         broadening
     )     
     
-    A_kω_Q_def = single_trion_arpes_signature_def(
+    @time A_kω_Q_def = single_trion_arpes_signature_def(
         ham, E_B, 
         ground_state_1s(ham, dielectric), 
         Q_point, 

### 2023-12-31 src/exciton-arpes-heatmap-thermalized.jl
update exciton code

@@ -1,0 +1,97 @@
+using ProgressMeter
+include("exciton-solver.jl")
+include("arpes.jl")
+
+# Collective exciton signature, with several possible exciton branches
+p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_v = 0.8184,
+    m_c = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7, # The value used in the literature: 0.7,
+    β = 10,
+    broadening(x) = exp(- σ^2 * x^2),
+    kx_points = LinRange(-1.2, 1.5, 100),
+    ky_points = LinRange(-1.2, 1.5, 100),
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(-1.2, 1.2, 100),
+    Q_points = k_points
+
+    ϵ_v(k) = - norm(k)^2 / 2m_v * inv_eV
+    ϵ_c(k) = E_g + norm(k - w)^2 / 2m_c * inv_eV
+
+    M = m_c + m_v
+    α = m_c / M
+    μ = m_c * m_v / (m_c + m_v)
+    a_ex = 0.529 * ϵ / μ
+    A_SQ(Q, k) = 1 / (1 + norm(k - (w + α * (Q - w)))^2 * a_ex^2 / 4)^2
+    E_SQ(Q) = E_g + E_B + (Q - w)^2/2M * inv_eV
+
+    ham = IndirectTwoBandModel2D(m_c, m_v, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+    Z_partition = sum(Q_points) do Q_point
+        exp(-β * (Q_point - [w, 0.0])' * (Q_point - [w, 0.0]) / 2M)
+    end
+    progress = Progress(length(Q_points)) 
+    A_kω_β = sum(Q_points) do Q_point
+        next!(progress)
+        A_kω_Q = single_exciton_arpes_signature_def(
+            ham, E_B, 
+            ground_state_1s(IndirectTwoBandExciton2D(ham, dielectric)), 
+            Q_point, 
+            ω_points, 
+            map(kx -> SVector{2, Float64}([kx, 0.0]), 
+                kx_points
+            ),
+            broadening
+        )   
[…]

### 2023-12-31 src/exciton-solver.jl

@@ -1,0 +1,58 @@
+using StaticArrays
+using ThreadTools
+
+include("two-band.jl")
+
+function single_exciton_arpes_signature_def(
+    ham::IndirectTwoBandModel2D,
+    E_B, 
+
+    A_SQ_k,
+    Q::SVector{2, Float64},
+    ω_points,
+    k_points_ARPES,
+    broadening,
+)
+    m_e = ham.m_e
+    m_h = ham.m_h
+    E_g = ham.E_g
+    w = ham.w
+    
+    M = m_e + m_h
+    E_SQ = inv_eV * norm(Q - w)^2 / 2M
+    
+    heatmap_points = Iterators.product(k_points_ARPES, ω_points)
+    A_kω = map(heatmap_points) do (k_e, ω)
+        k_h = Q - k_e
+
+        ϵ_v = - inv_eV * norm(k_h)^2 / 2m_h
+
+        δ_factor = broadening(ω - E_SQ - ϵ_v - E_B - E_g)
+        
+        k = k_e - m_e * Q / M - m_h * w / M
+        exciton_structure_factor = abs(A_SQ_k(k))^2
+        
+        δ_factor * exciton_structure_factor
+    end
+    
+    A_kω
+end
+
+struct IndirectTwoBandExciton2D 
+    ham::IndirectTwoBandModel2D
+    dielectric::Dielectric2D
+end
+
+function ground_state_1s(exciton_spec::IndirectTwoBandExciton2D)
+    ham = exciton_spec.ham
+    dielectric = exciton_spec.dielectric
+
+    m_c = ham.m_e
+    m_v = ham.m_h
+    ϵ = dielectric.ϵ
+
+    μ = m_c * m_v / (m_c + m_v)
+    a_ex = 0.529 * ϵ / μ
+
+    x -> ϕ_1s(x, a_ex)    
+end

### 2023-12-31 src/trion-arpes-heatmap-def.jl

@@ -3,7 +3,7 @@ include("trion-solver.jl")
 include("arpes.jl")
 
 # Finite momentum trion, slow version 
-@profview let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_h = 0.8184,
     m_e = 0.4794,
     E_g = 0.7830,
@@ -34,7 +34,7 @@ include("arpes.jl")
 
     A_kω_Q = single_trion_arpes_signature_def(
         ham, E_B, 
-        ground_state_1s(ham, dielectric), 
+        ground_state_1s_def(IndirectTwoBandTrion2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2023-12-31 src/trion-arpes-heatmap-optimized-one-Q.jl

@@ -34,7 +34,7 @@ include("arpes.jl")
 
     A_kω_Q = single_trion_arpes_signature_thread(
         ham, E_B, 
-        ground_state_1s(ham, dielectric), 
+        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2023-12-31 src/trion-arpes-heatmap-optimized-thermalized.jl

@@ -44,7 +44,7 @@ include("arpes.jl")
         next!(progress)
         A_kω_Q = single_trion_arpes_signature_thread(
             ham, E_B, 
-            ground_state_1s(ham, dielectric), 
+            ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
             Q_point, 
             k_points, ω_points, 
             map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2023-12-31 src/trion-arpes-optimization-test.jl

@@ -34,7 +34,7 @@ include("arpes.jl")
 
     @time A_kω_Q = single_trion_arpes_signature_thread(
         ham, E_B, 
-        ground_state_1s(ham, dielectric), 
+        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 
@@ -45,7 +45,7 @@ include("arpes.jl")
     
     @time A_kω_Q_def = single_trion_arpes_signature_def(
         ham, E_B, 
-        ground_state_1s(ham, dielectric), 
+        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2023-12-31 src/trion-solver.jl

@@ -1,20 +1,7 @@
 using StaticArrays
 using ThreadTools
 
-const fs = 1.5193    
-const inv_eV = 3.80998
-
-struct IndirectTwoBandModel2D
-    m_e :: Float64     # Conduction band effective mass
-    m_h :: Float64     # Valance band effective mass (positive)
-    E_g :: Float64     # Single particle band gap
-    w   :: SVector{2, Float64} # k_{valley} - k_{peak}
-end
-
-struct Dielectric2D 
-    ϵ :: Float64
-    # TODO
-end
+include("two-band.jl")
 
 
 function single_trion_arpes_signature_def(
@@ -100,10 +87,17 @@ function single_trion_arpes_signature_thread(
     A_kω
 end
 
-ϕ_1s_def(k::SVector{2, Float64}, a_ex::Float64) = 1 / (1 + norm(k)^2 * a_ex^2 / 4)^2
-@inline ϕ_1s(k::SVector{2, Float64}, a_ex::Float64) = @fastmath 1 / (1 + k' * k * a_ex^2 / 4)^2
 
-function ground_state_1s_def(ham::IndirectTwoBandModel2D, dielectric::Dielectric2D)
+
+struct IndirectTwoBandTrion2D 
+    ham::IndirectTwoBandModel2D
+    dielectric::Dielectric2D
+end
+
+function ground_state_1s_def(trion_spec::IndirectTwoBandTrion2D)
+    ham = trion_spec.ham
+    dielectric = trion_spec.dielectric
+
     m_c = ham.m_e
     m_v = ham.m_h
     ϵ = dielectric.ϵ
@@ -116,7 +110,10 @@ function ground_state_1s_def(ham::IndirectTwoBandModel2D, dielectric::Dielectric
     A_SQ_k1k2
 end
 
-function ground_state_1s(ham::IndirectTwoBandModel2D, dielectric::Dielectric2D)
+function ground_state_1s(trion_spec::IndirectTwoBandTrion2D)
+    ham = trion_spec.ham
+    dielectric = trion_spec.dielectric
+
     m_c = ham.m_e
     m_v = ham.m_h
     ϵ = dielectric.ϵ

### 2023-12-31 src/two-band.jl

@@ -1,0 +1,17 @@
+const fs = 1.5193    
+const inv_eV = 3.80998
+
+struct IndirectTwoBandModel2D
+    m_e :: Float64     # Conduction band effective mass
+    m_h :: Float64     # Valance band effective mass (positive)
+    E_g :: Float64     # Single particle band gap
+    w   :: SVector{2, Float64} # k_{valley} - k_{peak}
+end
+
+struct Dielectric2D 
+    ϵ :: Float64
+    # TODO
+end
+
+ϕ_1s_def(k::SVector{2, Float64}, a_ex::Float64) = 1 / (1 + norm(k)^2 * a_ex^2 / 4)^2
+@inline ϕ_1s(k::SVector{2, Float64}, a_ex::Float64) = @fastmath 1 / (1 + k' * k * a_ex^2 / 4)^2

### 2024-01-03 paper/main.tex
update exciton code
update paper

@@
 \newcommand*{\khi}[1]{\vb{k}_{\text{h#1}}}
+\newcommand*{\kh}{\vb{k}_{\text{h}}}
 \newcommand*{\re}{\vb{r}_{\text{e}}}
@@
 as the Hamiltonian in \cite{chang2021variationally}. 
+This can also be seen as generalization of the indirect band gap exciton two-band model to the trion;
+for the former we have \cite{rustagi2018photoemission} 
+\begin{equation}
+    H = \frac{(\vb{Q} - \vb{w})^2}{2M} + \Eg + \frac{\vb{k}^2}{2\mu} - V(\vb{r}), 
+\end{equation}
+where 
+\begin{equation}
+    M = \me + \me, \quad 
+    \ke = \frac{\me}{M} \vb{Q} + \vb{k} + \frac{\mh}{M} \vb{w}, \quad 
+    \kh = \frac{\mh}{M} \vb{Q} - \vb{k} - \frac{\mh}{M} \vb{w}.
+\end{equation}
 
@@
 and therefore the orbital part of the wave function is symmetric.
+
+It should be noted that since $\vb{k}$ is confined on the two-dimensional place, 
+so is $\vb{r}$, 
+which is the canonical coordinate corresponding to the crystal momentum $\vb{k}$
+instead of the free-space momentum, 
+and therefore differs from the free-space position;
+the wave function in terms of the free-space position does have $z$-directional distribution,
+but the wave function in terms of $\vb{r}$ does not.
+The spatial part of the wave function 
+therefore is to be solved on the 2D plane.
+On the other hand, electromagnetism in a two-dimensional system 
+is still inherently three-dimensional, 
+and the Coulomb force retains its inverse-square form,
+leading to the $\sim 1/r$ form of $V(r)$.
+
+\begin{equation}
+    \Psi(\vb{r}_1, \vb{r}_2) = 
+    \psi_{\text{1s,$a$}}(\vb{r}_1) 
+    \psi_{\text{1s,$b$}}(\vb{r}_2)
+    + \psi_{\text{1s,$b$}}(\vb{r}_1) 
+    \psi_{\text{1s,$a$}}(\vb{r}_2)
+\end{equation}
+
+\begin{equation}
+    \braket*{\text{1s,$a$}}{\text{1s,$b$}} = \frac{4ab}{(a+b)^2}.
+\end{equation}
+\begin{equation}
+    \mel*{\text{1s,$a$}}{-\laplacian}{\text{1s,$b$}} = \frac{4}{(a+b)^2}.
+\end{equation}
+
+\begin{equation}
+    \mel*{\text{1s, $a$}}{\frac{1}{r}}{\text{1s, $b$}} = \frac{4}{a+b}
+\end{equation}
 
@@
 The same procedure can be applied to other configurations of peaks and valleys.
+
+Variational wave function: TODO list 
+\begin{enumerate}
+    \item We should work in the real space, because the form of $V$ can be complicated.
+    \item The kinetic term should be evaluated analytically.
+    \item Normalization??? Why the factor is $1 / \sqrt{2}$?
+    \item The structure of the three-particle wave function.
+\end{enumerate}
 
@@
 
+It can be seen that the trion ARPES signature shows a significantly smaller variance with respect to the trion temperature
+than the exciton ARPES signature does.
+
 \printbibliography

### 2024-01-03 src/exciton-arpes-heatmap-thermalized.jl

@@ -2,7 +2,9 @@ using ProgressMeter
 include("exciton-solver.jl")
 include("arpes.jl")
 
-# Collective exciton signature, with several possible exciton branches
+# Interestingly, when I average over the 2D grid, 
+# the resulting signature loos less like the signature in Photoemission signature of excitons
+# I don't know: is the latter obtained by confining the problem on a 1D line?
 p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_v = 0.8184,
     m_c = 0.4794,
@@ -26,8 +28,6 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
     M = m_c + m_v
     α = m_c / M
     μ = m_c * m_v / (m_c + m_v)
-    a_ex = 0.529 * ϵ / μ
-    A_SQ(Q, k) = 1 / (1 + norm(k - (w + α * (Q - w)))^2 * a_ex^2 / 4)^2
     E_SQ(Q) = E_g + E_B + (Q - w)^2/2M * inv_eV
 
     ham = IndirectTwoBandModel2D(m_c, m_v, E_g, SVector{2, Float64}([w, 0.0]))

### 2024-01-03 src/exciton-solver.jl

@@ -52,7 +52,7 @@ function ground_state_1s(exciton_spec::IndirectTwoBandExciton2D)
     ϵ = dielectric.ϵ
 
     μ = m_c * m_v / (m_c + m_v)
-    a_ex = 0.529 * ϵ / μ
+    a_ex = a_Bohr * ϵ / μ
 
     x -> ϕ_1s(x, a_ex)    
 end

### 2024-01-03 src/trion-arpes-heatmap-optimized-thermalized.jl

@@ -10,9 +10,9 @@ include("arpes.jl")
     ϵ = 6.4, 
     E_B = -0.1,
     w = 0.7,
-    β = 10,
-    kx_points = LinRange(-1.2, 1.2, 100),
-    ky_points = LinRange(-1.2, 1.2, 100),
+    β = 1000,
+    kx_points = LinRange(-1.2, 1.5, 100),
+    ky_points = LinRange(-1.2, 1.5, 100),
     dkx = step(kx_points), 
     dky = step(ky_points),
     dk = dkx * dky,
@@ -86,6 +86,20 @@ include("arpes.jl")
         label = raw"Varying $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}} = \mathbf{k}_{\mathrm{2}} = 0$",
         linestyle = :dot,
         c = colorant"firebrick2")
+    # Dispersion relation when P_T is fixed to a constant 
+    plot!(p, kx_near_valley, 
+        ϵ_v.([[w, 0.0]] .- k_points_near_valley) 
+        .+ E_B .+ E_g,
+        #label = raw"fixed $\mathbf{P}_\mathrm{T}$, $\mathbf{k}_1 = 0$ or $\mathbf{k}_2 = 0$",
+        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}}$ or $\mathbf{k}_{\mathrm{2}}$ is $0$",
+        linestyle = :dot, 
+        c = colorant"aqua")
+    plot!(p, kx_near_valley, 
+        2 * ϵ_v.(([[w, 0.0]] .- k_points_near_valley) ./ 2) 
+        .+ E_B .+ E_g,
+        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}} = \mathbf{k}_{\mathrm{2}} $",
+        linestyle = :dot, 
+        c = colorant"deepskyblue2")
     
     ylims!(p, (-0.8, 1.8))
     xlims!(p, (-0.8, 1.2))

### 2024-01-03 src/trion-solver.jl

@@ -1,5 +1,6 @@
 using StaticArrays
 using ThreadTools
+using Symbolics
 
 include("two-band.jl")
 
@@ -103,7 +104,7 @@ function ground_state_1s_def(trion_spec::IndirectTwoBandTrion2D)
     ϵ = dielectric.ϵ
 
     μ = m_c * m_v / (m_c + m_v)
-    a_ex = 0.529 * ϵ / μ
+    a_ex = a_Bohr * ϵ / μ
 
     A_SQ_k1k2(k_1, k_2) = ϕ_1s_def(k_1, a_ex) * ϕ_1s_def(k_2, a_ex)
     
@@ -119,9 +120,25 @@ function ground_state_1s(trion_spec::IndirectTwoBandTrion2D)
     ϵ = dielectric.ϵ
 
     μ = m_c * m_v / (m_c + m_v)
-    a_ex = 0.529 * ϵ / μ
+    a_ex = a_Bohr * ϵ / μ
 
     @inline A_SQ_k1k2(k_1, k_2) = ϕ_1s(k_1, a_ex) * ϕ_1s(k_2, a_ex)
     
     A_SQ_k1k2
 end
+
+function ground_state_1sab_radii(trion_spec::IndirectTwoBandTrion2D)
+    ham = trion_spec.ham
+    dielectric = trion_spec.dielectric
+
+    m_c = ham.m_e
+    m_v = ham.m_h
+    ϵ = dielectric.ϵ
+
+    μ = m_c * m_v / (m_c + m_v)
+    a_ex = a_Bohr * ϵ / μ
+
+    ψ = ϕ_1s_def()  
+end
+
+@variables 

### 2024-01-03 src/two-band.jl

@@ -1,5 +1,6 @@
 const fs = 1.5193    
 const inv_eV = 3.80998
+const a_Bohr = 0.529
 
 struct IndirectTwoBandModel2D
     m_e :: Float64     # Conduction band effective mass

### 2024-01-03 src/real-space.jl
update real space code

@@ -1,0 +1,66 @@
+using StaticArrays
+using LinearAlgebra
+
+struct UniformGrid2D 
+    Lx::Float64
+    Ly::Float64
+    dx::Float64
+    dy::Float64
+    points::Vector{SVector{2, Float64}}
+end
+
+"""
+A two-dimensional grid.
+Lx and Ly are the size of the grid in the x and y directions
+(and therefore x_min ≈ -L/2, etc.)
+and Nx and Ny are the number of samples in the two directions.
+
+It's a good idea to set Nx and Ny to even numbers 
+if the (0, 0)  point is to be included into the grid.
+"""
+function UniformGrid2D(Lx, Ly, Nx, Ny)
+    dx = Lx / Nx 
+    dy = Ly / Ny
+    x_start = - Lx/2 + dx 
+    x_end   = Lx / 2
+    y_start = - Ly/2 + dy
+    y_end   = Ly / 2
+
+    points = reshape(
+        Iterators.product(x_start:dx:x_end, y_start:dy:y_end) |> collect |> permutedims, 
+        Nx * Ny
+    )
+    
+    UniformGrid2D(Lx, Ly, dx, dy, points)
+end
+
+ϕ_r_1s_def(r::SVector, a_ex::Float64) = sqrt(2 / (π * a_ex)) * exp(- norm(r) / a_ex)
+
+function overlap_numerical(ψ_1, ψ_2, grid::UniformGrid2D)
+    dA = grid.dx * grid.dy
+    dA * sum(grid.points) do pt
+        ψ_1(pt)' * ψ_2(pt) 
+    end
+end
+
+function coulomb_2D_numerical(ψ_1, ψ_2, grid::UniformGrid2D)
+    dA = grid.dx * grid.dy   
+    dA * sum(grid.points) do pt
+        r = norm(pt)
+        if r == 0.0
+            return 0.0 
+        end
+        ψ_1(pt)' * ψ_2(pt) * 1 / r 
+    end
+end
+
+function coulomb_twopoint_2D_numerical(ψ_1, ψ_2, grid::UniformGrid2D)
+    dA = grid.dx * grid.dy
+    dA^2 * sum(Iterators.product(grid.points, grid.points)) do (pt1, pt2) 
+        r12 = norm(pt1 - pt2)
+        if r12 == 0.0
+            return 0.0  
+        end
+        ψ_1(pt1)' * ψ_2(pt2) * 1 / r12
+    end
+end

### 2024-01-03 src/trion-solver.jl

@@ -1,7 +1,7 @@
 using StaticArrays
 using ThreadTools
-using Symbolics
 
+include("real-space.jl")
 include("two-band.jl")
 
 
@@ -140,5 +140,3 @@ function ground_state_1sab_radii(trion_spec::IndirectTwoBandTrion2D)
 
     ψ = ϕ_1s_def()  
 end
-
-@variables 

### 2024-01-03 src/two-band.jl

@@ -1,3 +1,5 @@
+using LinearAlgebra
+
 const fs = 1.5193    
 const inv_eV = 3.80998
 const a_Bohr = 0.529

### 2024-01-04 paper/main.tex
update real space code
minor updates

@@
 \begin{equation}
-    H =  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg 
-    + \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
-    + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2).
+    H =  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg +
+    \underbrace{
+        \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
+        + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
+    }_{H_{\text{internal}}}.
 \end{equation}
@@
 
+The 3D version of the problem is solved in \cite{chandrasekhar1944some,hogaasen2010two};
+
 \begin{equation}
@@
 
+Serious problem:
+\begin{equation}
+    \int \dd[2]{\vb{r}_1} \int \dd[2]{\vb{r}_2} \ee^{- r_1 / a} \frac{1}{\abs*{\vb{r}_1 - \vb{r}_2}} \ee^{- r_2 / b}
+\end{equation}
+One option might be to numerically solve the interaction part, 
+and use the analytic forms elsewhere.
+
 An immediate consequence of this fact is that
@@
     \item The structure of the three-particle wave function.
+    \item Also see the paper Two-electron atoms, ions, and molecules http://users.df.uba.ar/dmitnik/estructura3/articulostrabajo/twoelectronatomsionsmolecules.pdf
+    and https://articles.adsabs.harvard.edu/pdf/1944ApJ...100..176C
 \end{enumerate}

### 2024-01-04 src/real-space.jl

@@ -34,7 +34,7 @@ function UniformGrid2D(Lx, Ly, Nx, Ny)
     UniformGrid2D(Lx, Ly, dx, dy, points)
 end
 
-ϕ_r_1s_def(r::SVector, a_ex::Float64) = sqrt(2 / (π * a_ex)) * exp(- norm(r) / a_ex)
+ϕ_r_1s_def(r::SVector, a_ex::Float64) = sqrt(2 / (π * a_ex^2)) * exp(- norm(r) / a_ex)
 
 function overlap_numerical(ψ_1, ψ_2, grid::UniformGrid2D)
     dA = grid.dx * grid.dy

### 2024-01-04 test/real-space-test.jl

@@ -1,0 +1,54 @@
+include("../src/real-space.jl")
+using Plots
+
+
+let grid = UniformGrid2D(10, 10, 1000, 1000)
+    ψ_1s = x -> ϕ_r_1s_def(x, 1.5)
+    overlap_numerical(ψ_1s, ψ_1s, grid)
+end
+
+let Ns = [50, 100, 200, 500, 1000, 2000, 4000, 5000],
+    Ls = [10, 20, 30],
+    ψ_1s = x -> ϕ_r_1s_def(x, 2.8),
+    p = plot()
+    for L in Ls
+        overlap = map(Ns) do N
+            grid = UniformGrid2D(L, L, N, N)
+            overlap_numerical(ψ_1s, ψ_1s, grid)
+        end
+        plot!(p, log.(Ns), overlap, label = L, markershape = :circle)
+    end
+    p # It's expected that the values go to 1 when the parameters are large
+end
+
+let Ns = [50, 100, 200, 500, 1000, 2000, 4000, 5000],
+    Ls = [10, 20, 30],
+    a = 2.0,
+    ψ_1s = x -> ϕ_r_1s_def(x, a),
+    p = plot()
+    for L in Ls
+        E_coulomb = map(Ns) do N
+            grid = UniformGrid2D(L, L, N, N)
+            coulomb_2D_numerical(ψ_1s, ψ_1s, grid)
+        end
+        plot!(p, log.(Ns), E_coulomb, label = L, markershape = :circle)
+    end
+    hline!(p, [2 / a], label = "Analytic")
+    p # It's expected that the values go to 2/a when the parameters are large
+end
+
+let Ls = [10, 20, 30],
+    Ns = [50, 100, 200],    
+    a = 0.9,
+    ψ_1s = x -> ϕ_r_1s_def(x, a)
+    p = plot()
+    
+    for L in Ls
+        E_coulomb = map(Ns) do N
+            grid = UniformGrid2D(L, L, N, N)
+            coulomb_twopoint_2D_numerical(ψ_1s, ψ_1s, grid)
+        end
+        plot!(p, log.(Ns), E_coulomb, label = L, markershape = :circle)
+    end
+    p
+end

### 2024-01-04 test/trion-arpes-optimization-test.jl

@@ -1,6 +1,6 @@
 using ProgressMeter
-include("trion-solver.jl")
-include("arpes.jl")
+include("../src/trion-solver.jl")
+include("../src/arpes.jl")
 
 # Finite momentum trion, slow version 
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum

### 2024-01-12 src/exciton-arpes-heatmap-one-Q.jl
update data and report

@@ -12,7 +12,7 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
     ϵ = 6.4, 
     E_B = -0.1,
     w = 0.7, # The value used in the literature: 0.7,
-    β = 1,
+    β = 10,
     broadening(x) = exp(- σ^2 * x^2),
     kx_points = LinRange(-1.2, 1.5, 100),
     ky_points = LinRange(-1.2, 1.5, 100),
@@ -20,7 +20,8 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
         collect(Iterators.product(kx_points, ky_points))
     ),
     ω_points = LinRange(-1.2, 1.2, 100),
-    Q_points = k_points
+    Q = 0.7,
+    Q_point = SVector{2, Float64}([Q, 0.0])
 
     ϵ_v(k) = - norm(k)^2 / 2m_v * inv_eV
     ϵ_c(k) = E_g + norm(k - w)^2 / 2m_c * inv_eV
@@ -32,26 +33,19 @@ p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it shoul
 
     ham = IndirectTwoBandModel2D(m_c, m_v, E_g, SVector{2, Float64}([w, 0.0]))
     dielectric = Dielectric2D(ϵ)
-    Z_partition = sum(Q_points) do Q_point
-        exp(-β * (Q_point - [w, 0.0])' * (Q_point - [w, 0.0]) / 2M)
-    end
-    progress = Progress(length(Q_points)) 
-    A_kω_β = sum(Q_points) do Q_point
-        next!(progress)
-        A_kω_Q = single_exciton_arpes_signature_def(
-            ham, E_B, 
-            ground_state_1s(IndirectTwoBandExciton2D(ham, dielectric)), 
-            Q_point, 
-            ω_points, 
-            map(kx -> SVector{2, Float64}([kx, 0.0]), 
-                kx_points
-            ),
-            broadening
-        )   
-        exp(-β * (Q_point - [w, 0.0])' * (Q_point - [w, 0.0]) / 2M) * A_kω_Q / Z_partition
-    end  
+
+    A_kω_Q = single_exciton_arpes_signature_def(
+        ham, E_B, 
+        ground_state_1s(IndirectTwoBandExciton2D(ham, dielectric)), 
+        Q_point, 
+        ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
[…]

### 2024-01-12 src/trion-arpes-heatmap-optimized-one-Q-double-radii.jl

@@ -1,0 +1,111 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion, slow version 
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 1,
+    a = 10.3,
+    b = 25.2,
+    kx_points = LinRange(-1.2, 1.2, 100),
+    ky_points = LinRange(-1.2, 1.2, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.7,
+    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
+
+    A_kω_Q = single_trion_arpes_signature_thread(
+        ham, E_B, 
+        ground_state_1s_double_radii(IndirectTwoBandTrion2D(ham, dielectric), a, b), 
+        Q_point, 
+        k_points, ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )     
+    
+    p = heatmap(kx_points, ω_points, A_kω_Q', 
+        c = arpes_colormap(transparency_gradience),
+        colorbar = false, 
+        legend_foreground_color = :transparent,
+        legend = :topleft,
+        dpi = 500, 
+        grid = false, 
+        framestyle = :box,
[…]

### 2024-01-12 src/trion-arpes-heatmap-optimized-thermalized-double-radii.jl

@@ -1,0 +1,122 @@
+using ProgressMeter
+include("trion-solver.jl")
+include("arpes.jl")
+
+# Finite momentum trion, slow version 
+@time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.8184,
+    m_e = 0.4794,
+    E_g = 0.7830,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0.7,
+    β = 10,
+    a = 10.3,
+    b = 25.2,
+    kx_points = LinRange(-1.2, 1.5, 100),
+    ky_points = LinRange(-1.2, 1.5, 100),
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 1.5, 500),
+    P_Tx = 0.5,
+    Qx_points =  LinRange(w - 0.5, w + 0.5, 20),
+    Qy_points =  LinRange(- 0.5, 0.5, 20),
+    Q_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(Qx_points, Qy_points))
+    ) 
+
+    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
+    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    
+    M = 2m_h + m_e
+
+    ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
+
+    Z_partition = sum(Q_points) do Q_point
+        exp(-β * (Q_point - [w, 0.0])' * (Q_point - [w, 0.0]) / 2M)
+    end
+    progress = Progress(length(Q_points)) 
+    A_kω_β = sum(Q_points) do Q_point
+        next!(progress)
+        A_kω_Q = single_trion_arpes_signature_thread(
+            ham, E_B, 
+            ground_state_1s_double_radii(IndirectTwoBandTrion2D(ham, dielectric), a, b), 
+            Q_point, 
+            k_points, ω_points, 
[…]

### 2024-01-12 src/trion-solver.jl

@@ -127,7 +127,7 @@ function ground_state_1s(trion_spec::IndirectTwoBandTrion2D)
     A_SQ_k1k2
 end
 
-function ground_state_1sab_radii(trion_spec::IndirectTwoBandTrion2D)
+function ground_state_1s_double_radii(trion_spec::IndirectTwoBandTrion2D, a::Float64, b::Float64)
     ham = trion_spec.ham
     dielectric = trion_spec.dielectric
 
@@ -138,5 +138,11 @@ function ground_state_1sab_radii(trion_spec::IndirectTwoBandTrion2D)
     μ = m_c * m_v / (m_c + m_v)
     a_ex = a_Bohr * ϵ / μ
 
-    ψ = ϕ_1s_def()  
+    @inline ϕ_1s_a(k) = ϕ_1s(k, a)  
+    @inline ϕ_1s_b(k) = ϕ_1s(k, b)
+    @inline A_SQ_k1k2(k_1, k_2) = 1/sqrt(2) * (
+        ϕ_1s_a(k_1) * ϕ_1s_b(k_2) + ϕ_1s_a(k_2) * ϕ_1s_b(k_1)
+    )
+    
+    A_SQ_k1k2
 end

### 2024-02-06 paper/main.tex
update data and report
update report

@@
-\documentclass[hyperref, a4paper]{article}
+\documentclass[%
+ reprint,
+superscriptaddress,
+%groupedaddress,
+%unsortedaddress,
+%runinaddress,
+%frontmatterverbose, 
+%preprint,
+%preprintnumbers,
+%nofootinbib,
+%nobibnotes,
+%bibnotes,
+ amsmath,amssymb,
+ aps,
+%pra,
+prb,
+%rmp,
+%prstab,
+%prstper,
+%floatfix,
+]{revtex4-2}
 
 \usepackage{textgreek}
-\usepackage{geometry}
-\usepackage{titling}
-\usepackage{titlesec}
-\usepackage{footnote}
-\usepackage[colorinlistoftodos]{todonotes}
 \usepackage{booktabs}
@@
 \usepackage{graphicx}
-\usepackage{subcaption}
 \usepackage{physics}
@@
 \usepackage[most]{tcolorbox}
-\usepackage[backend=bibtex,sorting=none,doi=false,isbn=false,url=false]{biblatex}
-\addbibresource{trion.bib}
-\addbibresource{arpes.bib}
 
 \usepackage{prettyref}
-
-% Page style; to be removed when this article is placed in another template
-\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}
-\titlespacing{\paragraph}{0pt}{1pt}{10pt}[20pt]
-\setlength{\droptitle}{-5em}
 
@@
 \newcommand{\citetime}[1]{item \ref{#1} in \prettyref{sec:timeline}}
-\newcommand{\address}[1]{\href{#1}{\url{#1}}}
 \newcommand{\shortcode}[1]{\texttt{#1}}
@@
 \newcommand*{\ke}{\vb{k}_{\text{e}}}
+\newcommand*{\kei}[1]{\vb{k}_{\text{e#1}}}
 \newcommand*{\khi}[1]{\vb{k}_{\text{h#1}}}
 \newcommand*{\kh}{\vb{k}_{\text{h}}}
+\newcommand*{\kp}{\vb{k}_\parallel}
 \newcommand*{\re}{\vb{r}_{\text{e}}}
+\newcommand*{\rh}{\vb{r}_{\text{h}}}
 \newcommand*{\rhi}[1]{\vb{r}_{\text{h#1}}}
+\newcommand*{\rei}[1]{\vb{r}_{\text{e#1}}}
 \newcommand*{\me}{m_{\text{e}}}
@@
 
+\begin{document}
+
 \title{Trion ARPES signature}
-\author{Jinyuan Wu, someone else, Diana Y. Qiu}
-
-\begin{document}
+\author{Jinyuan Wu}
+\email{jinyuan.wu@yale.edu}
+\affiliation{Department of Mechanical Engineering and Materials Science, Yale University, New Haven, CT 06520}
+\author{Diana Y. Qiu}
+\email{diana.qiu@yale.edu}
+\affiliation{Department of Mechanical Engineering and Materials Science, Yale University, New Haven, CT 06520}
 
@@
 
-When the probe is much weaker than the pump 
+In most driven ARPES settings, the probe is much weaker than the pump 
 and does not itself influence the electronic structure,
-the most generic theory of driven ARPES is based on the  
-pumped single-electron lesser Green function, 
-to which the pump adds a self-energy correction,
+and the generic theory of ARPES can be formulated by 
+treating the pump as a non-equilibrium self-energy, 
+while the effect of the probe can be captured by the first order perturbation  
+and hence the final ARPES intensity is proportional to the electron dipole and 
+the pumped single-electron lesser Green function
 \cite{freericks2009theoretical,schuler2021theory};
 the theory reduced to the usual Fermi golden rule 
-when the pump itself is weak \cite{sobota2021angle,freericks2021two}.
+when the pump is turned off \cite{sobota2021angle,freericks2021two}.
 When the pump pulse is strong at its peak but has faded away at the probing time, 
@@
 pure-state first-order time dependent perturbation theory,
-which gives a Fermi golden rule-like final result \cite{rustagi2018photoemission}.
-Here we generalize the methodology in \cite{rustagi2018photoemission} 
-to all possible exciton states caused by the pump and not just excitons.
+which gives a Fermi golden rule-like expression of the probability of an electron being pumped out of the sample~\cite{rustagi2018photoemission}.
+Ref.~\cite{rustagi2018photoemission} is focused on excitons;
+here we generalize its methodology to all excitations and have  
+\begin{widetext}
+    \begin{equation}
+        I_{\vb{k}}(\omega, t) \propto \sum_{n_1, n_2} \sum_{n}
+        \rho_n (M^{fn_2}_{\vb{k} \vb{k}_\parallel})^* M^{fn_1}_{\vb{k} \vb{k}_\parallel} 
+        \int_{t_0}^t \dd{t_1} \int_{t_0}^t \dd{t_2}
+        \ee^{-\ii (E_n - \omega ) (t_1 - t_2)}
+        \mel{\Psi_n}{c^\dagger_{n_2 \vb{k}_\parallel} 
+        U(t_2, t_1) c_{n_1 \vb{k}_\parallel}}{\Psi_n} s(t_1) s(t_2),
+        \label{eq:exciton-response}
+    \end{equation}
+\end{widetext}
+where $t$ is the observation time, 
+$t_0$ is the starting time of the probe pulse,
+$s(t)$ is the envelope function of the probe,
+$\omega$ is the output electron energy shifted by the frequency of the probe pulse 
+and the work function of the sample, 
+$\vb{k}$ is the momentum of the outgoing electron 
+and $\vb{k}_\parallel$ is its projection on the two-dimensional plane,  
+$n_{1, 2}$ are electron band indices, 
+$n = (S, \vb{Q})$ is the index of exciton modes, 
+$\rho_n$ is the effective density matrix of excitons, 
+$\ket*{\Psi_n}$ is the many-body wave function of the exciton state at $t=t_0$,
+and $U(t_2, t_1)$ is the time evolution operator.
+The $c^\dagger U c \ee^{- \ii E_n t}$ 
+factor comes from the electron lesser Green function; 
+the $\ee^{- \ii E_n t}$ factor comes from the time evolution of the excitation states, 
+while the time evolution operator $U(t_2, t_1)$ 
+instructs the time evolution of the residue of the excitation after one electron is driven out; 
+time oscillation from the two factors together gives us the energy conservation relation.
+For three-dimensional systems, the momenta of $c$ and $c^\dagger$ in the $c^\dagger U c$ factor can be different \cite{freericks2009theoretical},  
+but for two-dimensional materials however, since by momentum conservation  
+the crystal momentum of the band electron that is driven out by the external field 
+can only be the parallel component of $\vb{k}$, and we refer it as $\vb{k}_\parallel$ in \eqref{eq:exciton-response}.
+As below we are mostly dealing with the ARPES spectrum of a single mode
+In principle, the $\mel{\Psi_n}{c^\dagger_{n_2 \vb{k}_\parallel} U(t_2, t_1) c_{n_1 \vb{k}_\parallel}}{\Psi_n}$ matrix can be non-diagonal with an excited initial state;
+frequently, however, because of energy conservation, the signature at given $(\omega, \vb{k})$
+can only come from a single band, and in this case $n_1 = n_2$.
 
@@
 \begin{equation}
-    H = \frac{(\ke - \vb{w})^2}{2\me} + \Eg 
-    + \frac{\khi{1}^2}{2\mh} + \frac{\khi{2}^2}{2\mh} 
-    + V(\rhi{1} - \rhi{2}) - V(\re - \rhi{1}) - V(\re - \rhi{2}).
+    \begin{aligned}
+        H &= \frac{(\ke - \vb{w})^2}{2\me} + \Eg 
+        + \frac{\khi{1}^2}{2\mh} + \frac{\khi{2}^2}{2\mh} \\
+        &\quad+ V(\rhi{1} - \rhi{2}) - V(\re - \rhi{1}) - V(\re - \rhi{2}).
+    \end{aligned}
     \label{eq:two-band-w}
 \end{equation}
-Here $\khi{i}$ is the crystal momentum of the quantum state containing hole $i$ only, 
+Here $\khi{$i$}$ is the crystal momentum of the quantum state containing hole $i$ only, 
 which is the opposite of the momentum of the corresponding electron state of the hole;
 similarly, $\vb{k}^2/2\mh$ is the opposite of the dispersion relation of the valence band top,
-and therefore $\mh > 0$.
+and therefore $\mh > 0$, 
+and under this Hamiltonian we treat holes as ordinary particles with positive masses and positive charges and ordinary momentum conservation relations without minus signs.
+
+The ARPES signature of trion mode $(S, \vb{P})$ comes solely from the electron in the trion.
+Therefore, the energy conservation relation hidden in \eqref{eq:exciton-response} means 
+the dispersion relation of the ARPES signature should always have the form of 
+\begin{equation}
+    \omega = E_n - \frac{\khi{1}^2}{2\mh} - \frac{\khi{2}^2}{2\mh}, 
+\end{equation}
+where $\khi{1, 2}$ are implicit functions of $\kp = \ke$ and $\vb{P}$.
+Unlike the case for excitons, however, 
+momentum conservation does not give a one-to-one mapping from $\ke$ and $\vb{P}$ to $\khi{1, 2}$; 
+we still need one additional constraint to pin down $\khi{1, 2}$, 
+which is to be given by the position of the trion wave function maximum when $\vb{P}$ and $\ke$ are given, 
+which, in turn, needs us to decompose \eqref{eq:two-band-w} into an external part that solely depends on the total momentum of the trion and an internal part that decides the structure of the trion wave function.
 
@@
 and the kinetic energy part of the \eqref{eq:two-band-w} now reads 
-\begin{equation}
+\[
     H_{\text{kin}} = \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg 
-    + \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh}
+    + \frac{\vb{k}_1^2 + \vb{k}_2^2}{2\mh} 
     - \frac{\vb{w}^2}{2M} + \frac{(\vb{k}_1 + \vb{k}_2 + \vb{w})^2}{2\me}.
-\end{equation}
+\]
 We however want to shift the definitions of $\vb{k}$ 
@@
 \begin{equation}
-    \vb{k}_{1,2} = - \ii \hbar \vb{r}_{1, 2} + \frac{\vb{w}}{1 + \me / \mu}, \quad 
+    \vb{k}_{1,2} = - \ii \hbar \partial_{\vb{r}_{1, 2}} + \frac{\vb{w}}{1 + \me / \mu}, \quad 
     \mu = \frac{\me \mh}{\me + \mh},
@@
 \begin{equation}
-    \ke = \frac{\me}{M} \vb{P} + \frac{2 \mh \vb{w}}{M} - \vb{k}_1 - \vb{k}_2, \quad 
-    \khi{1,2} = \vb{k}_{1,2} - \frac{\mh \vb{w}}{M} + \frac{\mh}{M} \vb{P},
+    \begin{aligned}
+        &\ke = \frac{\me}{M} \vb{P} + \frac{2 \mh \vb{w}}{M} - \vb{k}_1 - \vb{k}_2, \\ 
+        &\khi{1,2} = \vb{k}_{1,2} - \frac{\mh \vb{w}}{M} + \frac{\mh}{M} \vb{P},
+    \end{aligned}
 \end{equation}
@@
 \begin{equation}
-    H =  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg +
-    \underbrace{
-        \frac{\vb{k}_1^2}{2\mh} + \frac{\vb{k}_2^2}{2\mh} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
+    \begin{aligned}
+        H &=  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg, \\ 
+        &\ \  + \underbrace{
+            \frac{\vb{k}_1^2}{2\mu} + \frac{\vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
             + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
         }_{H_{\text{internal}}}.
+    \end{aligned}
+    \label{eq:internal-ham}
 \end{equation}
@@
 \begin{equation}
-    M = \me + \me, \quad 
-    \ke = \frac{\me}{M} \vb{Q} + \vb{k} + \frac{\mh}{M} \vb{w}, \quad 
-    \kh = \frac{\mh}{M} \vb{Q} - \vb{k} - \frac{\mh}{M} \vb{w}.
+    \begin{aligned}
+        &M = \me + \me, \\
+        &\ke = \frac{\me}{M} \vb{Q} + \vb{k} + \frac{\mh}{M} \vb{w}, \\
+        &\kh = \frac{\mh}{M} \vb{Q} - \vb{k} - \frac{\mh}{M} \vb{w}.
+    \end{aligned}
 \end{equation}
+
+The trion wave function therefore assumes the following form: 
+\begin{equation}
+    \psi_{S \vb{P}}(\vb{R}, \vb{r}_1, \vb{r}_2) = \ee^{\ii \vb{P} \cdot \vb{R}} \sum_{s_1, s_2} A_{s_1 s_2}^{S \vb{P}} \phi_{s_1}(\vb{r}_1) \phi_{s_2}(\vb{r}_2), 
+\end{equation}
+where $\phi_{s}(\vb{r})$ is an eigenfunction of the hydrogenic Hamiltonian $\frac{\vb{k}^2}{2\mu} - V(\vb{r})$, 
+and $A_{s_1 s_2}^{S \vb{P}}$ reflects the mixing of the modes by interactive Hamiltonian $V(\vb{r}_1 - \vb{r}_2)$ and $\vb{k}_1 \cdot \vb{k}_2 / 2\me$.
+Working in the momentum space we can also write 
+\begin{equation}
+    \psi_{S \vb{P}}(\vb{R}, \vb{p}_1, \vb{p}_2) = \ee^{\ii \vb{P} \cdot \vb{R}} \sum_{s_1, s_2} A_{s_1 s_2}^{S \vb{P}} \phi_{s_1}(\vb{p}_1) \phi_{s_2}(\vb{p}_2), 
+\end{equation}
+where $\phi_{s}(\vb{p})$ is the Fourier transform of $\phi_s(\vb{r})$.
+The additional constraint required to pin down the relation between $\khi{1,2}$ and $\ke, \vb{P}$ is maximization of $\psi_{S \vb{P}}(\vb{R}, \vb{p}_1, \vb{p}_2)$.
+
+\section{Two peak model}
+
+Similarly for 
+\begin{equation}
+    \begin{aligned}
+        H &= \frac{\ke^2}{2\me} + \Eg 
+        + \frac{\khi{1}^2}{2\mh} + \frac{(\khi{2} - \vb{w})^2}{2\mh} \\
+        &\ \ + V(\rhi{1} - \rhi{2}) - V(\re - \rhi{1}) - V(\re - \rhi{2}),
+    \end{aligned}
+\end{equation}
+by defining 
+\begin{equation}
+    \vb{k}_1 = - \ii \hbar \partial_{\vb{r}_{1}} + \frac{\mh}{M} \vb{w}, \quad 
+    \vb{k}_2 = - \ii \hbar \partial_{\vb{r}_{2}} - \frac{\me + \mh}{M} \vb{w},
+\end{equation}
+and hence 
+\begin{equation}
+    \begin{aligned}
+        &\ke = \frac{\me}{M} (\vb{P} - \vb{w}) - \vb{k}_1 - \vb{k}_2, \\ 
+        &\khi{1} = \vb{k}_1 - \frac{\mh}{M} \vb{w} + \frac{\mh}{M} \vb{P}, \\ 
+        &\khi{2} = \vb{k}_2 + \frac{\me + \mh}{M} \vb{w} + \frac{\mh}{M} \vb{P},
+    \end{aligned}
+\end{equation}
+we get exact the same Hamiltonian in \eqref{eq:internal-ham}.
+
+When $\vb{k}_1 = \vb{k}_2$, we have 
+\begin{equation}
+    \khi{1} = \frac{1}{2} (\vb{P} - \vb{w} - \ke), \quad 
+    \khi{2} = \frac{1}{2} (\vb{P} + \vb{w} - \ke), 
+\end{equation}
+and the dispersion relation is 
+\begin{equation}
+    \omega = E_{S\vb{P}} + \Eg + \frac{(\vb{P} - \vb{w} - \ke)^2}{4\mh}.
+\end{equation}
+
+In conclusion, the ARPES signature of any electron-hole-hole trion mode
+has two major differences from that of an exciton mode with a comparable momentum:
+the center and the curvature of the signature.
+
+\section{Two-electron case}
+
+The shifts of the center and the curvature of ARPES signatures are even more clearly 
+when the trion mode has two electrons and one hole.
+A typical setting of such a trion mode can be described by the following two-band model, 
+where the two electrons reside at different valleys:
+\begin{equation}
+    \begin{aligned}
+    H &= \frac{\kei{1}^2}{2\me} + \frac{(\kei{2} - \vb{w})^2}{2\me} + \frac{\kh^2}{2\mh} \\
+    &\quad + V(\rei{1} - \rei{2}) - V(\rh - \rei{1}) - V(\rh - \rei{2}).
+    \end{aligned}
+\end{equation}
+The decomposition of the Hamiltonian into internal and external terms 
+follows exactly the same procedure as in the two-peak model; 
+we only need to swap the positions of electrons and holes. 
+Therefore we get 
+\begin{equation}
+    \begin{aligned}
+        \vb{r}_1 &= \rei{1} - \rh, \quad 
+        \vb{r}_2 = \rei{2} - \rh, \\
+        \vb{k}_1 &= - \ii \partial_{\vb{r}_1} + \frac{\me}{M} \vb{w}, \quad 
+        \vb{k}_2 = - \ii \partial_{\vb{r}_2} - \frac{\me + \mh}{M} \vb{w}, 
+    \end{aligned}
+\end{equation}
+and conversely 
+\begin{equation}
+    \begin{aligned}
+        &\kh = \frac{\mh}{M} (\vb{P} - \vb{w}) - \vb{k}_1 - \vb{k}_2, \\
+        &\kei{1} = \vb{k}_1 - \frac{\me}{M} \vb{w} + \frac{\me}{M} \vb{P}, \\
+        &\kei{2} = \vb{k}_2 + \frac{\me + \mh}{M} \vb{w} + \frac{\me}{M} \vb{P} .
+    \end{aligned}
+\end{equation}
+Now when one electron is driven out by the probe, 
+it can be e1 or e2, so $\kp = \kei{1, 2}$, 
+momentum conservation blocks any block terms between the two valleys, 
+so we can evaluate the ARPES signature contributed by the two valleys separately.
+This drastically changes the dispersion relation of the ARPES signature, 
+but the structure of trion wave function is intact.
+Therefore, the curvature of the ARPES signature can still be found by either postulating $\vb{k}_1 = \vb{k}_2$ or by postulating that one of the momenta vanishes.
+When $\vb{k}_1 = \vb{k}_2$ and $\kp = \kei{1}$, we have 
+\begin{equation}
+    \kei{2} - \vb{w} = \kp, \quad \kh = \vb{P} - \vb{w} - 2 \kp, 
+\end{equation}
+and the dispersion relation is 
+\begin{equation}
+    \omega = E_{\text{bind}} - \frac{\kp^2}{2\me} - \frac{(2 \kp + \vb{w} - \vb{P})^2}{2\mh}, 
+\end{equation}
+while the effective mass is
 
@@
 
-\printbibliography
+\bibliography{arpes,trion}
 

### 2024-02-07 src/trion-arpes-heatmap-optimized-e1e2h1-double-radii.jl
update code structure and report

@@ -10,8 +10,9 @@ include("arpes.jl")
     ϵ = 6.4, 
     E_B = -0.1,
     w = 0.7,
-    β = 10,
-    broadening(x) = exp(- σ^2 * x^2),
+    β = 1,
+    a = 10.3,
+    b = 25.2,
     kx_points = LinRange(-1.2, 1.2, 100),
     ky_points = LinRange(-1.2, 1.2, 100),
     dkx = step(kx_points), 
@@ -21,21 +22,24 @@ include("arpes.jl")
         collect(Iterators.product(kx_points, ky_points))
     ),
     ω_points = LinRange(0, 1.5, 500),
-    P_Tx = 0.5,
-    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+    P_Tx = 0.7,
+    P_point  = SVector{2, Float64}([P_Tx, 0.0]) 
 
-    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
-    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
+    ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
+    ϵ_c(k)  =    norm(k)^2     / 2m_e * inv_eV + E_g
+    E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ E_g
     
     M = 2m_h + m_e
 
     ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
     dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = trionarpes_e2h1h1_def(
+    A_kω_Q = trionarpes_e1e2h1_thread(
         ham, E_B, 
-        phi1s_def(IndirectTwoBandMat2D(ham, dielectric)), 
-        Q_point, 
+        phi1sa1sb(IndirectTwoBandMat2D(ham, dielectric), a, b), 
+        P_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 
             kx_points
@@ -53,7 +57,11 @@ include("arpes.jl")
         framestyle = :box,
     )
 
-    plot!(p, kx_points, ϵ_v.(kx_points), 
[…]

### 2024-02-07 src/trion-arpes-heatmap-optimized-e1h1h2-double-radii.jl

@@ -10,8 +10,9 @@ include("arpes.jl")
     ϵ = 6.4, 
     E_B = -0.1,
     w = 0.7,
-    β = 10,
-    broadening(x) = exp(- σ^2 * x^2),
+    β = 1,
+    a = 10.3,
+    b = 25.2,
     kx_points = LinRange(-1.2, 1.2, 100),
     ky_points = LinRange(-1.2, 1.2, 100),
     dkx = step(kx_points), 
@@ -21,21 +22,24 @@ include("arpes.jl")
         collect(Iterators.product(kx_points, ky_points))
     ),
     ω_points = LinRange(0, 1.5, 500),
-    P_Tx = 0.5,
-    Q_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+    P_Tx = 0.7,
+    P_point  = SVector{2, Float64}([P_Tx, 0.0]) 
 
-    ϵ_v(k) =  - norm(k)^2 / 2m_h * inv_eV
-    ϵ_c(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
+    ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
+    ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
+    ϵ_c(k)  =    norm(k)^2     / 2m_e * inv_eV + E_g
+    E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ E_g
     
     M = 2m_h + m_e
 
     ham = IndirectTwoBandModel2D(m_e, m_h, E_g, SVector{2, Float64}([w, 0.0]))
     dielectric = Dielectric2D(ϵ)
+    broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = trionarpes_e2h1h1_def(
+    A_kω_Q = trionarpes_e1h1h2_thread(
         ham, E_B, 
-        phi1s_def(IndirectTwoBandMat2D(ham, dielectric)), 
-        Q_point, 
+        phi1sa1sb(IndirectTwoBandMat2D(ham, dielectric), a, b), 
+        P_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 
             kx_points
@@ -53,7 +57,11 @@ include("arpes.jl")
         framestyle = :box,
     )
 
-    plot!(p, kx_points, ϵ_v.(kx_points), 
[…]

### 2024-02-07 src/trion-arpes-heatmap-optimized-one-Q-double-radii.jl

@@ -34,9 +34,9 @@ include("arpes.jl")
     dielectric = Dielectric2D(ϵ)
     broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = single_trion_arpes_signature_thread(
+    A_kω_Q = trionarpes_e2h1h1_thread(
         ham, E_B, 
-        ground_state_1s_double_radii(IndirectTwoBandTrion2D(ham, dielectric), a, b), 
+        phi1sa1sb(IndirectTwoBandMat2D(ham, dielectric), a, b), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2024-02-07 src/trion-arpes-heatmap-optimized-one-Q.jl

@@ -32,9 +32,9 @@ include("arpes.jl")
     dielectric = Dielectric2D(ϵ)
     broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    A_kω_Q = single_trion_arpes_signature_thread(
+    A_kω_Q = trionarpes_e2h1h1_thread(
         ham, E_B, 
-        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
+        phi1s(IndirectTwoBandMat2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2024-02-07 src/trion-arpes-heatmap-optimized-thermalized-double-radii.jl

@@ -44,9 +44,9 @@ include("arpes.jl")
     progress = Progress(length(Q_points)) 
     A_kω_β = sum(Q_points) do Q_point
         next!(progress)
-        A_kω_Q = single_trion_arpes_signature_thread(
+        A_kω_Q = trionarpes_e2h1h1_thread(
             ham, E_B, 
-            ground_state_1s_double_radii(IndirectTwoBandTrion2D(ham, dielectric), a, b), 
+            phi1sa1sb(IndirectTwoBandMat2D(ham, dielectric), a, b), 
             Q_point, 
             k_points, ω_points, 
             map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2024-02-07 src/trion-arpes-heatmap-optimized-thermalized.jl

@@ -42,9 +42,9 @@ include("arpes.jl")
     progress = Progress(length(Q_points)) 
     A_kω_β = sum(Q_points) do Q_point
         next!(progress)
-        A_kω_Q = single_trion_arpes_signature_thread(
+        A_kω_Q = trionarpes_e2h1h1_thread(
             ham, E_B, 
-            ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
+            phi1s(IndirectTwoBandMat2D(ham, dielectric)), 
             Q_point, 
             k_points, ω_points, 
             map(kx -> SVector{2, Float64}([kx, 0.0]), 

### 2024-02-07 src/trion-solver.jl

@@ -5,7 +5,7 @@ include("real-space.jl")
 include("two-band.jl")
 
 
-function single_trion_arpes_signature_def(
+function trionarpes_e2h1h1_def(
     ham::IndirectTwoBandModel2D, 
     E_B,                                 # Binding energy in eV
     A_SQ_k1k2,                           # The internal part of the trion wave function  
@@ -46,9 +46,7 @@ function single_trion_arpes_signature_def(
     A_kω
 end
 
-single_trion_arpes_signature = single_trion_arpes_signature_def
-
-function single_trion_arpes_signature_thread(
+function trionarpes_e2h1h1_thread(
     ham::IndirectTwoBandModel2D, 
     E_B,                                 # Binding energy in eV
     A_SQ_k1k2,                           # The internal part of the trion wave function  
@@ -88,14 +86,109 @@ function single_trion_arpes_signature_thread(
     A_kω
 end
 
+function trionarpes_e1h1h2_thread(
+    ham::IndirectTwoBandModel2D, 
+    E_B,                                 # Binding energy in eV
+    A_SQ_k1k2,                           # The internal part of the trion wave function  
+    P_T::SVector{2, Float64},            # Total momentum of the trion  
+    k_points,       # The 1BZ sampling 
+    ω_points,       # The frequency sampling, in eV 
+    k_points_ARPES, # k points included in the ARPES plotting
+    broadening,     # Broadening function
+)
 
+    m_e = ham.m_e
+    m_h = ham.m_h
+    E_g = ham.E_g
+    M = 2m_h + m_e
+    w = ham.w
 
-struct IndirectTwoBandTrion2D 
-    ham::IndirectTwoBandModel2D
-    dielectric::Dielectric2D
+    E_S_PT = inv_eV * (P_T - w)' * (P_T - w) / 2M
+
+    heatmap_points = Iterators.product(k_points_ARPES, ω_points) 
+    
+    A_kω = tmap(heatmap_points) do (k_e, ω)
+        sum(k_points) do k_h1
+            k_h2 = P_T - k_e - k_h1
+            ϵ_v_kh1 = - inv_eV * k_h1' * k_h1 / 2m_h
[…]

### 2024-02-07 test/trion-arpes-optimization-test.jl

@@ -1,6 +1,9 @@
 using ProgressMeter
+using Test
 include("../src/trion-solver.jl")
 include("../src/arpes.jl")
+
+const TOL_SMALL = 1e-10
 
 # Finite momentum trion, slow version 
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
@@ -32,9 +35,9 @@ include("../src/arpes.jl")
     dielectric = Dielectric2D(ϵ)
     broadening(x) = @fastmath exp(- σ^2 * x^2)
 
-    @time A_kω_Q = single_trion_arpes_signature_thread(
+    @time A_kω_Q = trionarpes_e2h1h1_thread(
         ham, E_B, 
-        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
+        phi1s(IndirectTwoBandMat2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 
@@ -43,9 +46,9 @@ include("../src/arpes.jl")
         broadening
     )     
     
-    @time A_kω_Q_def = single_trion_arpes_signature_def(
+    @time A_kω_Q_def = trionarpes_e2h1h1_def(
         ham, E_B, 
-        ground_state_1s(IndirectTwoBandTrion2D(ham, dielectric)), 
+        phi1s(IndirectTwoBandMat2D(ham, dielectric)), 
         Q_point, 
         k_points, ω_points, 
         map(kx -> SVector{2, Float64}([kx, 0.0]), 
@@ -54,5 +57,5 @@ include("../src/arpes.jl")
         broadening
     )     
     
-    sum(abs.(A_kω_Q_def - A_kω_Q)) 
+    @test sum(abs.(A_kω_Q_def - A_kω_Q)) < TOL_SMALL
 end

### 2024-02-07 paper/main.tex
update trion arpes with e1e2h1; not 100percent sure

@@
 instructs the time evolution of the residue of the excitation after one electron is driven out; 
-time oscillation from the two factors together gives us the energy conservation relation.
+time oscillation from the two factors together gives us the energy conservation relation
+\begin{equation}
+    \omega = E_n - E_{\text{remain}} , 
+\end{equation}
+where $E_{\text{residue}}$ refers to the remaining energy of the system after
+one electron is driven out from $n_1, \kp$.
 For three-dimensional systems, the momenta of $c$ and $c^\dagger$ in the $c^\dagger U c$ factor can be different \cite{freericks2009theoretical},  

### 2024-02-07 src/trion-arpes-heatmap-optimized-e1e2h1-double-radii.jl

@@ -21,13 +21,14 @@ include("arpes.jl")
     k_points = map(t -> SVector{2, Float64}(collect(t)), 
         collect(Iterators.product(kx_points, ky_points))
     ),
-    ω_points = LinRange(0, 1.5, 500),
-    P_Tx = 0.7,
-    P_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+    ω_points = LinRange(-1, 1.5, 500),
+    P_x = 0.7,
+    P_point  = SVector{2, Float64}([P_x, 0.0]) 
 
     ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
     ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
-    ϵ_c(k)  =    norm(k)^2     / 2m_e * inv_eV + E_g
+    ϵ_c1(k) =    norm(k)^2     / 2m_e * inv_eV + E_g
+    ϵ_c2(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
     E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ E_g
     
     M = 2m_h + m_e
@@ -61,21 +62,22 @@ include("arpes.jl")
         label = "",
         linestyle = :dash,
         c = colorant"lightskyblue3")
-    plot!(p, kx_points, ϵ_v2.(kx_points), 
+    plot!(p, kx_points, ϵ_c1.(kx_points), 
         label = "",
         linestyle = :dash,
         c = colorant"lightskyblue3")
-    plot!(p, kx_points, ϵ_c.(kx_points), 
+    plot!(p, kx_points, ϵ_c2.(kx_points), 
         label = "",
         linestyle = :dash,
         c = colorant"lightskyblue3")
         
     kx_near_valley = LinRange(-0.7, 0.7, 100)
     k_points_near_valley = map(x -> [x, 0.0], kx_near_valley)
-    # k_1 when k_1 = k_2
-    k1_near_valley_equal   = 1/2 * ((P_Tx - w) * m_e / M .- kx_near_valley)
+    # k_1 when e1 is out  
+    k1_near_valley_e1 = kx_near_valley .+ m_e / M * (w - P_x) 
[…]

### 2024-02-07 src/trion-solver.jl

@@ -150,10 +150,10 @@ function trionarpes_e1e2h1_thread(
     A_kω_1 = tmap(heatmap_points) do (k_e1, ω)
         sum(k_points) do k_e2
             k_h = P_T - k_e1 - k_e2
-            ϵ_c2 =   inv_eV * (k_e2 - w)' * (k_e2 - w) / 2m_e
-            ϵ_v  = - inv_eV * k_h' * k_h / 2m_h
+            ϵ_c2 =  inv_eV * (k_e2 - w)' * (k_e2 - w) / 2m_e
+            ϵ_v  =  inv_eV * k_h' * k_h / 2m_h
 
-            δ_factor = broadening(ω - E_S_PT - ϵ_c2 - ϵ_v - E_B - E_g)
+            δ_factor = broadening(ω - E_S_PT + ϵ_c2 + ϵ_v - E_B)
             
             k_1 = k_e1 + m_e / M * w - m_e / M * P_T
             k_2 = k_e2 - (m_e + m_h) / M * w - m_e / M * P_T
@@ -166,10 +166,10 @@ function trionarpes_e1e2h1_thread(
     A_kω_2 = tmap(heatmap_points) do (k_e2, ω)
         sum(k_points) do k_e1
             k_h = P_T - k_e1 - k_e2
-            ϵ_c1 =   inv_eV * k_e1' * k_e1 / 2m_e
-            ϵ_v  = - inv_eV * k_h' * k_h / 2m_h
+            ϵ_c1 = inv_eV * k_e1' * k_e1 / 2m_e
+            ϵ_v  = inv_eV * k_h' * k_h / 2m_h
 
-            δ_factor = broadening(ω - E_S_PT - ϵ_c1 - ϵ_v - E_B - E_g)
+            δ_factor = broadening(ω - E_S_PT + ϵ_c1 + ϵ_v - E_B)
             
             k_1 = k_e1 + m_e / M * w - m_e / M * P_T
             k_2 = k_e2 - (m_e + m_h) / M * w - m_e / M * P_T

### 2024-02-14 paper/main.tex
update eeh trion

@@
 \begin{equation}
-    \omega = E_n - E_{\text{remain}} , 
+    \omega = E_n - E_{\text{residue}} , 
 \end{equation}
@@
     \begin{aligned}
-        H &=  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg, \\ 
+        H &=  \frac{(\vb{P} - \vb{w})^2}{2 M} + \Eg \\ 
         &\ \  + \underbrace{
-            \frac{\vb{k}_1^2}{2\mu} + \frac{\vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
+            \frac{\vb{k}_1^2 + \vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
             + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
@@
 We find that the internal part of the Hamiltonian has the same form 
-as the Hamiltonian in \cite{chang2021variationally}. 
+as the trion Hamiltonian in \cite{chang2021variationally}, 
+which includes no indirect band gap.
 This can also be seen as generalization of the indirect band gap exciton two-band model to the trion;
-for the former we have \cite{rustagi2018photoemission} 
+for the exciton case we have \cite{rustagi2018photoemission} 
 \begin{equation}
@@
 
-The trion wave function therefore assumes the following form: 
+After diagonalization of \eqref{eq:internal-ham}, we get a trion spectrum with the form of 
+\begin{equation}
+    E_{S \vb{P}} = \frac{(\vb{P} - \vb{w})^2}{2M} + \Eg + E^{\text{b}}_{S},
+\end{equation}
+where $E_S^{\text{b}}$ comes from diagonalization of $H_{\text{internal}}$.
+The trion wave function coming from diagonalization of  $H_{\text{internal}}$ assumes the following form: 
 \begin{equation}
@@
     \begin{aligned}
-    H &= \frac{\kei{1}^2}{2\me} + \frac{(\kei{2} - \vb{w})^2}{2\me} + \frac{\kh^2}{2\mh} \\
+    H &= \frac{\kei{1}^2}{2\me} + \frac{(\kei{2} - \vb{w})^2}{2\me} + \frac{\kh^2}{2\mh} + 2 \Eg \\
     &\quad + V(\rei{1} - \rei{2}) - V(\rh - \rei{1}) - V(\rh - \rei{2}).
@@
 we only need to swap the positions of electrons and holes. 
-Therefore we get 
+Therefore we find that under 
 \begin{equation}
@@
 \end{equation}
-and conversely 
+or conversely 
 \begin{equation}
@@
         &\kei{1} = \vb{k}_1 - \frac{\me}{M} \vb{w} + \frac{\me}{M} \vb{P}, \\
-        &\kei{2} = \vb{k}_2 + \frac{\me + \mh}{M} \vb{w} + \frac{\me}{M} \vb{P} .
+        &\kei{2} = \vb{k}_2 + \frac{\me + \mh}{M} \vb{w} + \frac{\me}{M} \vb{P} ,
     \end{aligned}
 \end{equation}
+the Hamiltonian of the two-electron-one-hole trion is reduced to 
+% See src/two-electron.nb
+\begin{equation}
+    \begin{aligned}
+        H &=  \frac{(\vb{P} - \vb{w})^2}{2 M} + 2 \Eg, \\ 
+        &\ \  + \underbrace{
+            \frac{\vb{k}_1^2}{2\mu} + \frac{\vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\mh}
+            + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
+        }_{H_{\text{internal}}},
+    \end{aligned}
+\end{equation}
+a form resembling \eqref{eq:internal-ham}, 
+the only differences being the coefficient of the $\vb{k}_1 \cdot \vb{k}_2$ term
+and the $2\Eg$ term due to the fact that there are two electrons now.
 Now when one electron is driven out by the probe, 
-it can be e1 or e2, so $\kp = \kei{1, 2}$, 
+it can be e1 or e2, so $\kp = \kei{1,2}$, 
 momentum conservation blocks any block terms between the two valleys, 
@@
 but the structure of trion wave function is intact.
+The most remarkable difference between the ARPES signature of a negative trion 
+and that of a positive trion
+is the former has a vertical ``tail'': 
+when $\kp = \kei{1}$ is chosen, $\vb{k}_2$ is allow to move freely, 
+and it continuously adjusts the energy of the remaining of the trion after e1 is driven out, 
+which means the ARPES signature will have a long, vertical component 
+that does not coincide with any dispersion relations. % TODO: rework the wording
+
 Therefore, the curvature of the ARPES signature can still be found by either postulating $\vb{k}_1 = \vb{k}_2$ or by postulating that one of the momenta vanishes.
@@
 \begin{equation}
-    \omega = E_{\text{bind}} - \frac{\kp^2}{2\me} - \frac{(2 \kp + \vb{w} - \vb{P})^2}{2\mh}, 
+    \omega = \Eg + E_{\text{b}} + \frac{\vb{P}^2}{2M} - \frac{\kp^2}{2\me} - \frac{(2 \kp + \vb{w} - \vb{P})^2}{2\mh}, 
 \end{equation}

### 2024-02-14 src/trion-arpes-heatmap-optimized-e1e2h1-double-radii.jl

@@ -2,7 +2,6 @@ using ProgressMeter
 include("trion-solver.jl")
 include("arpes.jl")
 
-# Finite momentum trion, slow version 
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_h = 0.8184,
     m_e = 0.4794,
@@ -29,7 +28,7 @@ include("arpes.jl")
     ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
     ϵ_c1(k) =    norm(k)^2     / 2m_e * inv_eV + E_g
     ϵ_c2(k) =    norm(k - w)^2 / 2m_e * inv_eV + E_g
-    E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ E_g
+    E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ 2E_g
     
     M = 2m_h + m_e
 
@@ -100,7 +99,7 @@ include("arpes.jl")
         - ϵ_c2.(k1_near_valley_e1 .+ (m_e + m_h) / M * w .+ m_e / M * P_x) 
         .+ ϵ_v1.(m_h / M * (P_x - w) .- 2k1_near_valley_e1)
         .+ E_SP(P_x), 
-        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}}$ or $\mathbf{k}_{\mathrm{2}}$ is $0$",
+        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}} = \mathbf{k}_{\mathrm{2}} $",
         linestyle = :dot, 
         c = colorant"aqua")
     # e1 is driven out, leaving e2 and h; 
@@ -109,7 +108,7 @@ include("arpes.jl")
         - ϵ_c2.((m_e + m_h) / M * w .+ m_e / M * P_x) 
         .+ ϵ_v1.(m_h / M * (P_x - w) .- k1_near_valley_e1)
         .+ E_SP(P_x), 
-        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}} = \mathbf{k}_{\mathrm{2}} $",
+        label = raw"Fixed $\mathbf{P}$, $\mathbf{k}_{\mathrm{1}}$ or $\mathbf{k}_{\mathrm{2}}$ is $0$",
         linestyle = :dot, 
[…]

### 2024-02-14 src/trion-solver.jl

@@ -153,7 +153,7 @@ function trionarpes_e1e2h1_thread(
             ϵ_c2 =  inv_eV * (k_e2 - w)' * (k_e2 - w) / 2m_e
             ϵ_v  =  inv_eV * k_h' * k_h / 2m_h
 
-            δ_factor = broadening(ω - E_S_PT + ϵ_c2 + ϵ_v - E_B)
+            δ_factor = broadening(ω - E_S_PT - E_B - E_g + ϵ_c2 + ϵ_v)
             
             k_1 = k_e1 + m_e / M * w - m_e / M * P_T
             k_2 = k_e2 - (m_e + m_h) / M * w - m_e / M * P_T
@@ -169,7 +169,7 @@ function trionarpes_e1e2h1_thread(
             ϵ_c1 = inv_eV * k_e1' * k_e1 / 2m_e
             ϵ_v  = inv_eV * k_h' * k_h / 2m_h
 
-            δ_factor = broadening(ω - E_S_PT + ϵ_c1 + ϵ_v - E_B)
+            δ_factor = broadening(ω - E_S_PT - E_B - E_g + ϵ_c1 + ϵ_v)
             
             k_1 = k_e1 + m_e / M * w - m_e / M * P_T
             k_2 = k_e2 - (m_e + m_h) / M * w - m_e / M * P_T

### 2024-03-28 src/arpes-makie.jl
minor update

@@ -1,0 +1,29 @@
+using CairoMakie
+using LinearAlgebra
+using Colors
+using ColorSchemes
+using LaTeXStrings
+
+function arpes_colormap(gradient)
+    rainbow = cgrad(:rainbow)
+    starting_region = gradient(first(rainbow), steps = 50) 
+    middle_region = rainbow[LinRange(0, 1, 500)]
+    end_region = reverse(gradient(last(rainbow), steps = 400))
+    
+    vcat(starting_region, middle_region, end_region)
+end
+
+function transparency_gradience(color; steps = 100)
+    α = alpha(color)
+    r = red(color)
+    b = blue(color)
+    g = green(color)
+    
+    map(LinRange(0.0, α, steps)) do α_new
+        RGBA(r, g, b, α_new)
+    end
+end
+
+function white_gradience(color; steps = 100)
+    ColorScheme([colorant"white", color])[LinRange(0, 1, steps)]
+end

### 2024-03-28 src/trion-arpes-heatmap-optimized-e1e2h1-double-radii-makie.jl

@@ -1,26 +1,26 @@
 using ProgressMeter
 include("trion-solver.jl")
-include("arpes.jl")
+include("arpes-makie.jl")
 
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
-    m_h = 0.8184,
-    m_e = 0.4794,
-    E_g = 0.7830,
+    m_h = 0.21,
+    m_e = 0.37,
+    E_g = 2.84,
     ϵ = 6.4, 
     E_B = -0.1,
     w = 0.7,
     β = 1,
     a = 10.3,
     b = 25.2,
-    kx_points = LinRange(-1.2, 1.2, 100),
-    ky_points = LinRange(-1.2, 1.2, 100),
+    kx_points = LinRange(-1.2, 1.2, 200),
+    ky_points = LinRange(-1.2, 1.2, 200),
     dkx = step(kx_points), 
     dky = step(ky_points),
     dk = dkx * dky,
     k_points = map(t -> SVector{2, Float64}(collect(t)), 
         collect(Iterators.product(kx_points, ky_points))
     ),
-    ω_points = LinRange(-1, 1.5, 500),
+    ω_points = LinRange(-1, 3, 500),
     P_x = 0.7,
     P_point  = SVector{2, Float64}([P_x, 0.0]) 
 
@@ -47,28 +47,38 @@ include("arpes.jl")
         broadening
     )     
     
-    p = heatmap(kx_points, ω_points, A_kω_Q', 
-        c = arpes_colormap(transparency_gradience),
-        colorbar = false, 
-        legend_foreground_color = :transparent,
-        legend = :topleft,
-        dpi = 500, 
-        grid = false, 
-        framestyle = :box,
+    fig = Figure()
+    ax_heatmap = Axis(fig[1, 1], 
+        xlabel = L"$k$ (Å)", 
+        ylabel = L"$ω$ (eV)",
+        xlabelsize = 20,
+        ylabelsize = 20,
+        xticklabelsize = 18,
+        yticklabelsize = 18,
     )
 
-    plot!(p, kx_points, ϵ_v1.(kx_points), 
-        label = "",
+    heatmap!(ax_heatmap, kx_points, ω_points, A_kω_Q, 
[…]

### 2024-03-29 src/exciton-arpes-heatmap-one-Q-makie.jl
more formal figures

@@ -1,0 +1,104 @@
+using ProgressMeter
+include("exciton-solver.jl")
+include("arpes-makie.jl")
+
+# Interestingly, when I average over the 2D grid, 
+# the resulting signature loos less like the signature in Photoemission signature of excitons
+# I don't know: is the latter obtained by confining the problem on a 1D line?
+p = let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_v = 0.21,
+    m_c = 0.37,
+    E_g = 2.84,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    w = 0,#4π / (3 * 3.144817974),
+    β = 10,
+    broadening(x) = exp(- σ^2 * x^2),
+    kx_points = LinRange(-1.2, 1.5, 100),
+    ky_points = LinRange(-1.2, 1.5, 100),
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    ω_points = LinRange(0, 3, 500),
+    hexagonal_edge = 4π / (3 * 3.144817974),
+    Q = 0.2 * hexagonal_edge,
+    Q_point = SVector{2, Float64}([Q, 0.0])
+
+    ϵ_v(k) = - norm(k)^2 / 2m_v * inv_eV
+    ϵ_c(k) = E_g + norm(k - w)^2 / 2m_c * inv_eV
+
+    M = m_c + m_v
+    α = m_c / M
+    μ = m_c * m_v / (m_c + m_v)
+    E_SQ(Q) = E_g + E_B + (Q - w)^2/2M * inv_eV
+
+    ham = IndirectTwoBandModel2D(m_c, m_v, E_g, SVector{2, Float64}([w, 0.0]))
+    dielectric = Dielectric2D(ϵ)
+
+    A_kω_Q = single_exciton_arpes_signature_def(
+        ham, E_B, 
+        ground_state_1s(IndirectTwoBandExciton2D(ham, dielectric)), 
+        Q_point, 
+        ω_points, 
+        map(kx -> SVector{2, Float64}([kx, 0.0]), 
+            kx_points
+        ),
+        broadening
+    )   
+    
+    fig = Figure()
+    ax_heatmap = Axis(fig[1, 1], 
+        xlabel = L"$k$ (Å)", 
+        ylabel = L"$ω$ (eV)",
+        xlabelsize = 20,
[…]

### 2024-03-29 src/trion-arpes-heatmap-optimized-e1e2h1-double-radii-makie.jl

@@ -8,20 +8,21 @@ include("arpes-makie.jl")
     E_g = 2.84,
     ϵ = 6.4, 
     E_B = -0.1,
-    w = 0.7,
+    hexagonal_edge = 4π / (3 * 3.144817974),
+    w = hexagonal_edge,
     β = 1,
     a = 10.3,
     b = 25.2,
-    kx_points = LinRange(-1.2, 1.2, 200),
-    ky_points = LinRange(-1.2, 1.2, 200),
+    kx_points = LinRange(-0.35, 1.7, 200),
+    ky_points = LinRange(-0.35, 1.7, 200),
     dkx = step(kx_points), 
     dky = step(ky_points),
     dk = dkx * dky,
     k_points = map(t -> SVector{2, Float64}(collect(t)), 
         collect(Iterators.product(kx_points, ky_points))
     ),
-    ω_points = LinRange(-1, 3, 500),
-    P_x = 0.7,
+    ω_points = LinRange(0, 4, 500),
+    P_x = hexagonal_edge,
     P_point  = SVector{2, Float64}([P_x, 0.0]) 
 
     ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
@@ -80,7 +81,7 @@ include("arpes-makie.jl")
         colorrange = (0, 10),
     )
         
-    kx_near_valley = LinRange(-0.7, 0.7, 100)
+    kx_near_valley = LinRange(minimum(kx_points), maximum(kx_points), 100)
     k_points_near_valley = map(x -> [x, 0.0], kx_near_valley)
     # k_1 when e1 is out  
     k1_near_valley_e1 = kx_near_valley .+ m_e / M * (w - P_x) 
@@ -154,13 +155,24 @@ include("arpes-makie.jl")
         colorrange = (0, 10),
     )
 
-    axislegend(ax_heatmap, backgroundcolor = RGBA(1, 1, 1, 0.5), framecolor = RGBA(1, 1, 1, 0.5), labelsize = 18)
+    axislegend(ax_heatmap, 
+        backgroundcolor = RGBA(1, 1, 1, 0.5), 
+        framecolor = RGBA(1, 1, 1, 0.5), 
+        labelsize = 18, 
+        position = :rb,
+    )
 
     ylims!(ax_heatmap, (-0.8, 4))
-    xlims!(ax_heatmap, (-0.8, 1.2))
+    xlims!(ax_heatmap, (minimum(kx_points), maximum(kx_points)))
[…]

### 2024-03-29 src/trion-arpes-heatmap-optimized-e1h1h2-double-radii-makie.jl

@@ -1,15 +1,16 @@
 using ProgressMeter
 include("trion-solver.jl")
-include("arpes.jl")
+include("arpes-makie.jl")
 
 # Finite momentum trion, slow version 
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
-    m_h = 0.8184,
-    m_e = 0.4794,
-    E_g = 0.7830,
+    m_h = 0.21,
+    m_e = 0.37,
+    E_g = 2.84,
     ϵ = 6.4, 
     E_B = -0.1,
-    w = 0.7,
+    hexagonal_edge = 4π / (3 * 3.144817974),
+    w = hexagonal_edge,
     β = 1,
     a = 10.3,
     b = 25.2,
@@ -21,8 +22,8 @@ include("arpes.jl")
     k_points = map(t -> SVector{2, Float64}(collect(t)), 
         collect(Iterators.product(kx_points, ky_points))
     ),
-    ω_points = LinRange(0, 1.5, 500),
-    P_Tx = 0.9,
+    ω_points = LinRange(0, 3, 500),
+    P_Tx = 1.2 * w,
     P_point  = SVector{2, Float64}([P_Tx, 0.0]) 
 
     ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
@@ -47,28 +48,38 @@ include("arpes.jl")
         broadening
     )     
     
-    p = heatmap(kx_points, ω_points, A_kω_Q', 
-        c = arpes_colormap(transparency_gradience),
-        colorbar = false, 
-        legend_foreground_color = :transparent,
-        legend = :topleft,
-        dpi = 500, 
-        grid = false, 
-        framestyle = :box,
+    fig = Figure()
+    ax_heatmap = Axis(fig[1, 1], 
+        xlabel = L"$k$ (Å)", 
+        ylabel = L"$ω$ (eV)",
+        xlabelsize = 20,
+        ylabelsize = 20,
+        xticklabelsize = 18,
+        yticklabelsize = 18,
     )
 
-    plot!(p, kx_points, ϵ_v1.(kx_points), 
-        label = "",
+    heatmap!(ax_heatmap, kx_points, ω_points, A_kω_Q,
+        colormap = arpes_colormap(transparency_gradience),
+    )
+    
[…]

### 2024-04-28 paper/main.tex
minor update
more formal figures
update julia env and code

@@
 \usepackage{booktabs}
+\usepackage{subcaption}
 \usepackage{array}
@@
 \author{Jinyuan Wu}
-\email{jinyuan.wu@yale.edu}
 \affiliation{Department of Mechanical Engineering and Materials Science, Yale University, New Haven, CT 06520}
@@
 
-\section{Two-band model for trion}
+\section{Two-band model for a positive trion}
 
@@
 to the indirect band gap case.
-We start with the following example.
+We start with the following Hamiltonian of a negative trion, 
+i.e. a trion that contains one electron and two holes (ehh).
+The effective Hamiltonian reads
 \begin{equation}
@@
         &\ \  + \underbrace{
-            \frac{\vb{k}_1^2 + \vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\me}
+            \frac{\vb{k}_1^2 + \vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{\me}
             + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
@@
 
+\begin{figure*}
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/exciton-Q-0.2-w-0.0-makie.png}      
+        \subcaption{}
+    \end{subfigure}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{figs/trion-e1h1h2-P-1.2-w-1.0-makie.png}
+        \subcaption{}
+    \end{subfigure}
+    \caption{The simulated ARPES spectra of the exciton (a) and the ehh trion (b). 
+    Effective masses and the band gap are from Ref. \cite{qiu2013optical}.
+    The curvatures of the two signals are different, 
+    and so are the center of them.}
+\end{figure*}
+
 Similarly for 
@@
 \end{equation}
+We note that the effective mass of this dispersion relation is $2 \mh$ instead of $\mh$,
+meaning that the ARPES signature does not have a well-defined curvature corresponding to an effective mass of $\mh$.
+Moreover, the center of the ARPES signature is predicted to be 
+\begin{equation}
+    \ke = \frac{\me}{M} (\vb{P} - \vb{w}),
+\end{equation}
+different from
 
@@
         &\ \  + \underbrace{
-            \frac{\vb{k}_1^2}{2\mu} + \frac{\vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{2\mh}
+            \frac{\vb{k}_1^2}{2\mu} + \frac{\vb{k}_2^2}{2\mu} + \frac{\vb{k}_1 \cdot \vb{k}_2}{\mh}
             + V(\vb{r}_1 - \vb{r}_2) - V(\vb{r}_1) - V(\vb{r}_2)
@@
 \end{equation}
-while the effective mass is
+while the effective mass is  TODO: an analytic expression giving the shape of the signature is still elusive
+
+\begin{figure}
+    \includegraphics[width=0.4\textwidth]{figs/trion-e1e2h1-P-1.0-w-1.0-makie.png}
+    \caption{The simulated ARPES spectrum of the eeh trion. Effective masses and the band gap are from Ref.~\cite{qiu2013optical}.}
+\end{figure}
 
@@
 
-\bibliography{arpes,trion}
+\bibliography{arpes,trion,materials}
 

### 2024-04-28 src/arpes-makie.jl

@@ -1,4 +1,5 @@
 using CairoMakie
+using MakieTeX
 using LinearAlgebra
 using Colors
 using ColorSchemes

### 2024-04-28 src/trion-arpes-heatmap-optimized-e1h1h2-double-radii-makie-decomposed.jl

@@ -1,0 +1,217 @@
+include("trion-solver.jl")
+include("arpes-makie.jl")
+using Printf
+
+##
+
+σ = 20fs
+m_h = 0.21
+m_e = 0.37
+E_g = 2.84
+ϵ = 6.4
+E_B = -0.1
+hexagonal_edge = 4π / (3 * 3.144817974)
+w = hexagonal_edge
+β = 1
+a = 10.3
+b = 25.2
+kx_points = LinRange(-0.7, 1.7, 200)
+ky_points = LinRange(-0.7, 1.7, 200)
+dkx = step(kx_points)
+dky = step(ky_points)
+dk = dkx * dky
+k_points = map(t -> SVector{2, Float64}(collect(t)),
+    collect(Iterators.product(kx_points, ky_points))
+)
+ω_points = LinRange(0, 3, 500)
+P_Tx = 1.2 * w
+P_T = SVector{2, Float64}([P_Tx, 0.0]) 
+
+ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
+ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
+ϵ_c(k)  =    norm(k)^2     / 2m_e * inv_eV + E_g
+E_SP(P) = inv_eV * norm(P .- w)^2 / 2M .+ E_B .+ E_g
+
+M = 2m_h + m_e
+
+
+
+A_kω_Q = let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
+    m_h = 0.21,
+    m_e = 0.37,
+    E_g = 2.84,
+    ϵ = 6.4, 
+    E_B = -0.1,
+    hexagonal_edge = 4π / (3 * 3.144817974),
+    w = hexagonal_edge,
+    β = 1,
+    a = 10.3,
+    b = 25.2,
+    dkx = step(kx_points), 
+    dky = step(ky_points),
+    dk = dkx * dky,
+    k_points = map(t -> SVector{2, Float64}(collect(t)), 
+        collect(Iterators.product(kx_points, ky_points))
+    ),
+    P_Tx = 1.2 * w,
+    P_point  = SVector{2, Float64}([P_Tx, 0.0]) 
+
+    ϵ_v2(k) =  - norm(k - w)^2 / 2m_h * inv_eV
+    ϵ_v1(k) =  - norm(k)^2     / 2m_h * inv_eV
[…]

### 2024-04-28 src/trion-arpes-heatmap-optimized-e1h1h2-double-radii-makie.jl

@@ -2,7 +2,7 @@ using ProgressMeter
 include("trion-solver.jl")
 include("arpes-makie.jl")
 
-# Finite momentum trion, slow version 
+# Finite momentum trion
 @time let σ = 20fs, # Note that here σ tells us the width of the pulse; it should be *large* to produce δ-function like ARPES spectrum
     m_h = 0.21,
     m_e = 0.37,
